---
title: "KZFPs"
output:
  html_document:
    toc: true
    toc_float: true
date: "2026-07-20"
---

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

# KZFP Expression

## Setting up the environment

### Loading libraries
```{r Enviroment Set Up, }
#loading libraries
library("tidyverse")
library("rmarkdown")
library("readxl")
library("readr")
library("ggrepel")
library("ggfortify")
library("patchwork")
library("eulerr")
library("org.Hs.eg.db")
library("AnnotationDbi")
library("ggVennDiagram")
library("ggrastr")
library("edgeR")
library("limma")
library("ggsignif")
library("dplyr")
library("purrr")
library("ggplot2")
library("tidyr")
library("AnnotationDbi")
library("org.Hs.eg.db")
library("ggalluvial")

# BiocManager::install("ggalluvial")

```

### Defining Functions
```{r unctions}
KZFP_enrichment <- function(comparisons, TopTable, KZFPs_list) {
  
  results <- lapply(comparisons, function(group) {
    
    test_genes <- TopTable %>%
      filter(Comparisons == group) %>%
      distinct(Symbol)
    
    deg_genes <- TopTable %>%
      filter(Comparisons == group,
             adj.P.Val < 0.05) %>%
      distinct(Symbol)
    
    not_deg_genes <- setdiff(test_genes$Symbol, deg_genes$Symbol)
    
    tab <- matrix(
      c(
        sum(deg_genes$Symbol %in% KZFPs_list),
        sum(deg_genes$Symbol %notin% KZFPs_list),
        sum(not_deg_genes %in% KZFPs_list),
        sum(not_deg_genes %notin% KZFPs_list)
      ),
      nrow = 2,
      byrow = TRUE
    )
    
    fisher <- fisher.test(tab)
    
    data.frame(
      Comparison = group,
      KZFP_DEG = tab[1,1],
      NonKZFP_DEG = tab[1,2],
      KZFP_not_DEG = tab[2,1],
      NonKZFP_not_DEG = tab[2,2],
      Odds_Ratio = fisher$estimate,
      Fisher_p = fisher$p.value
    )
  })
  
  do.call(rbind, results)
}

#KZFP_enrichment_fate_function
Fate_KZFP_enrichment <- function(comparisons, TopTable_Fate, KZFPs_list) {
  
  results <- lapply(comparisons, function(group) {
    
    test_genes <- TopTable_Fate %>%
      filter(Comparisons == group) %>%
      distinct(Symbol, Fate)
    
    
    fate_tests <- list(
      Specific = "Specific",
      Shared_partial = "Shared_partial",
      Shared_all = "Shared_all"
    )
    
    
    do.call(rbind, lapply(names(fate_tests), function(fate_name) {
      
      fate_genes <- test_genes %>%
        filter(Fate == fate_tests[[fate_name]]) %>%
        pull(Symbol)
      
      non_fate_genes <- test_genes %>%
        filter(Fate != fate_tests[[fate_name]]) %>%
        pull(Symbol)
      
      
      tab <- matrix(
        c(
          sum(fate_genes %in% KZFPs_list),
          sum(fate_genes %notin% KZFPs_list),
          sum(non_fate_genes %in% KZFPs_list),
          sum(non_fate_genes %notin% KZFPs_list)
        ),
        nrow = 2,
        byrow = TRUE
      )
      
      fisher <- fisher.test(tab)
      
      data.frame(
        Comparison = group,
        Fate = fate_name,
        KZFP_Fate = tab[1,1],
        NonKZFP_Fate = tab[1,2],
        KZFP_nonFate = tab[2,1],
        NonKZFP_nonFate = tab[2,2],
        Odds_Ratio = fisher$estimate,
        Fisher_p = fisher$p.value
      )
      
    }))
  })
  
  do.call(rbind, results)
}


# Fisher Test Odds Ratio plotting

plot_OR_cluster <- function(results, cluster_to_plot, ylim = c(0, 2)) {

  plot_data <- results %>%
    filter(Cluster %in% cluster_to_plot) %>%
    mutate(
      label = ifelse(FDR < 0.05, "*", "")
    )

  ggplot(plot_data, aes(Cluster, Odds_Ratio)) +
    geom_point(size = 3) +
    geom_text(
      aes(label = label),
      vjust = -1
    ) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    coord_cartesian(ylim = ylim) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      x = "CorMotif Cluster",
      y = "Odds Ratio"
    )
}

#FischerTest_OddsRation_plotting

plot_OR <- function(results, group_to_plot, ylim = c(-2, 2)) {

  plot_data <- results %>%
    filter(Comparison_Group %in% group_to_plot) %>%
    mutate(
      label = ifelse(FDR < 0.05, "*", ""),
      log2_OR = log2(Odds_Ratio)
    )

  ggplot(plot_data, aes(Comparison, Odds_Ratio)) +
    geom_point(size = 3) +
    geom_text(
      aes(label = label),
      vjust = -1
    ) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    coord_cartesian(ylim = ylim) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

##Cluster Enrichemnt 
KZFP_cluster_enrichment_Unassigned <- function(clusters, TopTable, KZFPs_list) {
  
  results <- lapply(clusters, function(cluster) {
    
    cluster_genes <- TopTable %>%
      filter(Cluster == cluster) %>%
      distinct(Symbol) %>%
      pull(Symbol)
    
    unassigned_genes <- TopTable %>%
      filter(Cluster == "Unassigned") %>%
      distinct(Symbol) %>%
      pull(Symbol)
    
    tab <- matrix(
      c(
        sum(cluster_genes %in% KZFPs_list),
        sum(!cluster_genes %in% KZFPs_list),
        sum(unassigned_genes %in% KZFPs_list),
        sum(!unassigned_genes %in% KZFPs_list)
      ),
      nrow = 2,
      byrow = TRUE
    )
    
    fisher <- fisher.test(tab)
    
    data.frame(
      Cluster = cluster,
      KZFP_Cluster = tab[1,1],
      NonKZFP_Cluster = tab[1,2],
      KZFP_Unassigned = tab[2,1],
      NonKZFP_Unassigned = tab[2,2],
      Odds_Ratio = fisher$estimate,
      Fisher_p = fisher$p.value
    )
  })
  
  do.call(rbind, results)
}

##Cluster Enrichemnt 
cluster_enrichment_NoResponse <- function(clusters, TopTable, list) {
  
  results <- lapply(clusters, function(cluster) {
    
    cluster_genes <- TopTable %>%
      filter(Cluster == cluster) %>%
      distinct(Symbol) %>%
      pull(Symbol)
    
    unassigned_genes <- TopTable %>%
      filter(Cluster == "cluster1") %>%
      distinct(Symbol) %>%
      pull(Symbol)
    
    tab <- matrix(
      c(
        sum(cluster_genes %in% list),
        sum(!cluster_genes %in% list),
        sum(unassigned_genes %in% list),
        sum(!unassigned_genes %in% list)
      ),
      nrow = 2,
      byrow = TRUE
    )
    
    fisher <- fisher.test(tab)
    
    data.frame(
      Cluster = cluster,
      List_Cluster = tab[1,1],
      NonList_Cluster = tab[1,2],
      List_Cluster1 = tab[2,1],
      NonList_Cluster1 = tab[2,2],
      Odds_Ratio = fisher$estimate,
      Fisher_p = fisher$p.value
    )
  })
  
  do.call(rbind, results)
}


#FischerTest_fate_OddsRation_plotting

plot_Fate_OR <- function(results, ylim = c(0, 2)) {

  plot_data <- results %>%
    mutate(
      label = ifelse(FDR < 0.05, "*", ""),
      log2_OR = log2(Odds_Ratio)
    )

  ggplot(plot_data, aes(x = Comparison, y = Odds_Ratio)) +
    geom_point(size = 3) +
    geom_text(
      aes(label = label),
      vjust = -1
    ) +
    geom_hline(
      yintercept = 1,
      linetype = "dashed"
    ) +
    facet_wrap(~Fate) +
    coord_cartesian(ylim = ylim) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      x = "HvC developmental comparison",
      y = "Odds Ratio",
      title = "KZFP enrichment by HvC DEG fate"
    )
}

#Volcano_Plot_Function

generate_volcano_plot <- function(toptable, title) {
  
  # #check for entrezid 
  # if(!"Entrez_ID" %in% colnames(toptable)) stop("Entrez_ID col not present")
  # 
  #make significance labels
  
  toptable <- toptable %>% 
    mutate(Significance = case_when(
      logFC > 0 & adj.P.Val < 0.05 ~ "Upregulated",
      logFC < 0 & adj.P.Val < 0.05 ~ "Downregulated",
      TRUE  ~ "Not Significant"
    ))
  
  #factor significance
  toptable$Significance <- factor(
    toptable$Significance, 
    levels = c("Upregulated", 
               "Not Significant", 
               "Downregulated")
  )
  
  #count genes in each category
  upgenes <- sum(toptable$Significance == "Upregulated")
  nsgenes <- sum(toptable$Significance == "Not Significant")
  downgenes <- sum(toptable$Significance == "Downregulated")
  
  
  #labels for legend
  legend_lab <- c(
    paste0("Upregulated: ", upgenes),
    paste0("Not Significant: ", nsgenes),
    paste0("Downregulated: ", downgenes)
  )
  
  #colors 
  color_map <- c("Upregulated" = "blue",
                 "Not Significant" = "grey30",
                 "Downregulated" = "red")
  
  #generate volcano plots
  p <- ggplot(toptable, aes(x = logFC, 
                            y = -log10(P.Value),
                            color = Significance)) +
    geom_point_rast(alpha = 0.8, size = 3) +
    scale_color_manual(values = color_map, 
                       labels = legend_lab,
                       breaks = c("Upregulated",
                                  "Not Significant", 
                                  "Downregulated"))  +
    xlim(-20,20) +
    labs(title = title, 
         x = expression("log"[2]*"FC"),
         y = expression("-log"[10]*"p-value")) +
    # theme_custom() +
    theme(legend.position = "right")
  
  return(p)
}



#FC_Plot_Function_sig_trajectory_No4

plot_cluster_fc_sig_trajectory_No4 <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(logFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$logFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = logFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("logFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG FC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(-2, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}


#absFC_Plot_Function_sig_trajectory_No4

plot_cluster_absfc_sig_trajectory_No4 <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(absFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$absFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = absFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("absFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG absFC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(0, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}


#get_directional_genes

get_fc_directions <- function(traj_label, coef_name, comparison_label,
                              top_table_all, fit_obj, p_cutoff = 0.05, gene_set = NULL) {

  # ---------------------------
  # trajectory FC data
  # ---------------------------
  df_fc <- top_table_all %>%
    dplyr::filter(
      Comparisons %in% c(
        paste0("H_", comparison_label),
        paste0("C_", comparison_label)
      )
    )

  df_dir <- df_fc %>%
    dplyr::mutate(Species = ifelse(grepl("^H_", Comparisons), "H", "C")) %>%
    dplyr::group_by(Gene, Species) %>%
    dplyr::summarise(logFC = mean(logFC, na.rm = TRUE), .groups = "drop") %>%
    tidyr::pivot_wider(names_from = Species, values_from = logFC) %>%
    dplyr::filter(!is.na(H), !is.na(C)) %>%
    dplyr::distinct(Gene, .keep_all = TRUE)

  # ---------------------------
  # KEY FILTER: DEGs in BOTH species
  # ---------------------------
  if (!is.null(gene_set)) {
    df_dir <- df_dir %>%
      dplyr::filter(Gene %in% gene_set)
  }

  # ---------------------------
  # direction classification
  # ---------------------------
  df_combo <- df_dir %>%
    dplyr::mutate(
      direction = dplyr::case_when(
        H > 0 & C < 0 ~ "H_pos_C_neg",
        H < 0 & C > 0 ~ "H_neg_C_pos",
        H > 0 & C > 0 ~ "Both_pos",
        H < 0 & C < 0 ~ "Both_neg",
        TRUE ~ "Other"
      ),
      direction_group = dplyr::case_when(
        H * C > 0 ~ "Same_direction",
        H * C < 0 ~ "Opposite_direction",
        TRUE ~ "Other"
      )
    )

  # ---------------------------
  # output
  # ---------------------------
  list(
    all = df_combo,
    H_pos_C_neg = dplyr::filter(df_combo, direction == "H_pos_C_neg"),
    H_neg_C_pos = dplyr::filter(df_combo, direction == "H_neg_C_pos"),
    Both_pos = dplyr::filter(df_combo, direction == "Both_pos"),
    Both_neg = dplyr::filter(df_combo, direction == "Both_neg"),

    Same_direction = dplyr::filter(df_combo, direction_group == "Same_direction"),
    Opposite_direction = dplyr::filter(df_combo, direction_group == "Opposite_direction"),

    Other = dplyr::filter(df_combo, direction == "Other"),

    n_sig = length(gene_set)
  )
}

#get_directional_genes_HvC1

get_fc_directions <- function(traj_label, coef_name, comparison_label,
                                 top_table_all, fit_obj, p_cutoff = 0.05) {

  # ---------------------------
  # DE genes
  # ---------------------------
  tt <- topTable(fit_obj, coef = coef_name, number = Inf)

  sig_tt <- tt %>%
    dplyr::filter(adj.P.Val < p_cutoff) %>%
    tibble::rownames_to_column("Gene") %>%
    dplyr::distinct(Gene, .keep_all = TRUE)

  # ---------------------------
  # HvC FC data
  # ---------------------------
  df_fc <- top_table_all %>%
    dplyr::filter(Comparisons == comparison_label) %>%
    dplyr::distinct(Gene, .keep_all = TRUE)

  # ---------------------------
  # classify direction
  # ---------------------------
  df_combo <- df_fc %>%
  dplyr::inner_join(sig_tt, by = "Gene") %>%
  dplyr::mutate(
    direction = dplyr::case_when(
      logFC.x > 0 ~ "Increased",
      logFC.x < 0 ~ "Decreased",
      TRUE ~ "No_difference"
    ),
    magnitude = abs(logFC.x)
  )

  # ---------------------------
  # output
  # ---------------------------
  list(
    all = df_combo,
    Increased = dplyr::filter(df_combo, direction == "Increased"),
    Decreased = dplyr::filter(df_combo, direction == "Decreased"),
    n_sig = nrow(sig_tt)
  )
}
```

### Data Upload
```{r TopTable_Symbols}
# TopTable <- readRDS("data/DGE/TopTable_logcpm_G1_SpTp_All_NoRep.RDS")
# 
# # Extract Ensembl IDs for every row
# ens_ids <- sub("\\..*$", "", TopTable$Gene)
# 
# # Make a lookup table
# symbol_map <- mapIds(
#     org.Hs.eg.db,
#     keys = unique(ens_ids),
#     column = "SYMBOL",
#     keytype = "ENSEMBL",
#     multiVals = "first"
# )
# 
# TopTable$Symbol <- unname(symbol_map[ens_ids])
# 
# saveRDS(TopTable,"data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
# 
# TopTable <- readRDS("data/DGE/TopTable_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
# 
# # Extract Ensembl IDs for every row
# ens_ids <- sub("\\..*$", "", TopTable$Gene)
# 
# # Make a lookup table
# symbol_map <- mapIds(
#     org.Hs.eg.db,
#     keys = unique(ens_ids),
#     column = "SYMBOL",
#     keytype = "ENSEMBL",
#     multiVals = "first"
# )
# 
# TopTable$Symbol <- unname(symbol_map[ens_ids])
# 
# saveRDS(TopTable,"data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

# TF_df <- read.delim("C:/Users/jdhurley/Downloads/protein_class_Transcription (1).tsv")
# saveRDS(TF_df,"data/TEs/TF_df.RDS")


# library(readxl)
# KZFP_metadata_Hardy2023 <- read_excel("C:/Users/jdhurley/Downloads/KZFP_metadata_Hardy2023.xlsx", 
#     sheet = "Primary_targets")
# 
# saveRDS(KZFP_metadata_Hardy2023,"data/TEs/KZFPs_ages.RDS")
```

```{r KZFPs}
KZFPs <- read_excel("data/TEs/KZFPs_list.xlsx")

# saveRDS(KZFPs,"data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

# saveRDS(KZFPs_list,"data/TEs/KZFPs_list.RDS")

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

intersect(TopTable$Symbol,KZFPs_list)
```


## Enrichment Analysis

### All

```{r KZFP_enrichment,fig.height=4, fig.width=4}

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")
TopTable <- readRDS(
  "data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS"
)

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

# Existing enrichment results
comparisons <- unique(TopTable$Comparisons)

KZFP_results <- KZFP_enrichment(
  comparisons,
  TopTable,
  TF_list
)

# --------------------------------------------------
# InteractionAtLeastOne
# --------------------------------------------------

Inter_02_genes <- unique(RNA_sets$Sig_Sp_inter_Traj02$DEGs$Gene)
Inter_25_genes <- unique(RNA_sets$Sig_Sp_inter_Traj25$DEGs$Gene)
Inter_515_genes <- unique(RNA_sets$Sig_Sp_inter_Traj515$DEGs$Gene)
Inter_1530_genes <- unique(RNA_sets$Sig_Sp_inter_Traj1530$DEGs$Gene)

InteractionAtLeastOne <- unique(c(
  Inter_02_genes,
  Inter_25_genes,
  Inter_515_genes,
  Inter_1530_genes
))

# Universe = all genes represented in TopTable
test_genes <- unique(TopTable$Symbol)

# Convert InteractionAtLeastOne Ensembl IDs to gene symbols
InteractionAtLeastOne_symbols <- TopTable %>%
  filter(Gene %in% InteractionAtLeastOne) %>%
  pull(Symbol) %>%
  unique()

# InteractionAtLeastOne genes are DEGs by definition
deg_genes <- intersect(
  InteractionAtLeastOne_symbols,
  test_genes
)

# Everything else in the universe is non-DEG
not_deg_genes <- setdiff(
  test_genes,
  deg_genes
)

tab <- matrix(
  c(
    sum(deg_genes %in% KZFPs_list),
    sum(deg_genes %notin% KZFPs_list),
    sum(not_deg_genes %in% KZFPs_list),
    sum(not_deg_genes %notin% KZFPs_list)
  ),
  nrow = 2,
  byrow = TRUE
)

fisher <- fisher.test(tab)

InteractionAtLeastOne_result <- data.frame(
  Comparison = "InteractionAtLeastOne",
  KZFP_DEG = tab[1,1],
  NonKZFP_DEG = tab[1,2],
  KZFP_not_DEG = tab[2,1],
  NonKZFP_not_DEG = tab[2,2],
  Odds_Ratio = fisher$estimate,
  Fisher_p = fisher$p.value
)

# Add to existing results
KZFP_results <- bind_rows(
  KZFP_results,
  InteractionAtLeastOne_result
)

# BH correction across the complete set of comparisons
KZFP_results$FDR <- p.adjust(
  KZFP_results$Fisher_p,
  method = "BH"
)

comparison_order <- c(
  "HvC_Day0",
  "HvC_Day2",
  "HvC_Day5",
  "HvC_Day15",
  "HvC_Day30",
  "H_Day02",
  "H_Day25",
  "H_Day515",
  "H_Day1530",
  "C_Day02",
  "C_Day25",
  "C_Day515",
  "C_Day1530",
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530",
  "InteractionAtLeastOne"
)


KZFP_results <- KZFP_results %>%
  mutate(
    Comparison = factor(
      Comparison,
      levels = comparison_order
    ),
    Comparison_Group = case_when(
      Comparison == "InteractionAtLeastOne" ~ "Interaction",
      grepl("^HvC", Comparison) ~ "Species",
      grepl("inter", Comparison) ~ "Interaction",
      grepl("^[HC]_Day", Comparison) ~ "Trajectory",
      TRUE ~ "Other"
    ),
    Significant = FDR < 0.05
  )

KZFP_results

plot_OR(KZFP_results, "Species", ylim = c(0, 2))
plot_OR(KZFP_results, "Trajectory", ylim = c(0, 2))
plot_OR(KZFP_results, "Interaction", ylim = c(0, 2))

```

#### Log fold change
```{r KZFP_Species_FC}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

comparisons <- unique(TopTable$Comparisons)

comparison_order <- c(
  "HvC_Day0",
  "HvC_Day2",
  "HvC_Day5",
  "HvC_Day15",
  "HvC_Day30",
  "H_Day02",
  "H_Day25",
  "H_Day515",
  "H_Day1530",
  "C_Day02",
  "C_Day25",
  "C_Day515",
  "C_Day1530",
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530"
)

KZFP_all <- TopTable %>%
  filter(
    Comparisons %in% comparison_order,
    Symbol %in% KZFPs_list
  ) %>%
  mutate(
    Comparisons = factor(
      Comparisons,
      levels = comparison_order
    )
  )

ggplot(KZFP_all, aes(Comparisons, logFC)) +
  geom_boxplot() +
  ggtitle("Logfc of all expressed KZFPs") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  coord_cartesian(ylim = c(-2, 2)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

KZFP_all$absfc <- abs(KZFP_all$logFC)

ggplot(KZFP_all, aes(Comparisons, absfc)) +
  geom_boxplot() +
  ggtitle("Absolute Logfc of all expressed KZFPs") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  coord_cartesian(ylim = c(0, 3)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

saveRDS(KZFP_all,"data/TEs/KZFPs_Toptable.RDS")

KZFP_FC <- TopTable %>%
  filter(
    Comparisons %in% comparisons,
    Symbol %in% KZFPs_list
  ) %>%
  dplyr::select(Comparisons, Symbol, logFC, adj.P.Val)

KZFP_DEG <- TopTable %>%
  filter(
    Comparisons %in% comparisons,
    Symbol %in% KZFPs_list,
    adj.P.Val < 0.05
  ) %>%
  dplyr::select(Comparisons, Symbol, logFC, adj.P.Val)

KZFP_DEG %>%
  filter(grepl("^HvC", Comparisons)) %>%
  arrange(Comparisons, desc(logFC))

KZFP_DEG_direction <- KZFP_DEG %>%
  mutate(
    Direction = case_when(
      
      # Species contrasts
      grepl("^HvC", Comparisons) & logFC > 0 ~ "Higher in Human",
      grepl("^HvC", Comparisons) & logFC < 0 ~ "Higher in Chimp",
      
      # Human trajectory contrasts
      grepl("^H_Day", Comparisons) & logFC > 0 ~ "Human increased",
      grepl("^H_Day", Comparisons) & logFC < 0 ~ "Human decreased",
      
      # Chimp trajectory contrasts
      grepl("^C_Day", Comparisons) & logFC > 0 ~ "Chimp increased",
      grepl("^C_Day", Comparisons) & logFC < 0 ~ "Chimp decreased",
      
      # Species x trajectory interactions
      grepl("^Sig_Sp_inter", Comparisons) & logFC > 0 ~ "Human-Chimp difference increased",
      grepl("^Sig_Sp_inter", Comparisons) & logFC < 0 ~ "Human-Chimp difference decreased",
      
      TRUE ~ "Unknown"
    )
  ) %>%
  dplyr::count(Comparisons, Direction)
```


```{r KZFP_Species_FC_lineplot}
KZFP_all <-readRDS("data/TEs/KZFPs_Toptable.RDS")

library(dplyr)
library(ggplot2)

# Split the existing KZFP data into the three comparison types
KZFP_HvC <- KZFP_all %>%
  filter(grepl("^HvC_", Comparisons))

KZFP_Traj <- KZFP_all %>%
  filter(grepl("^[HC]_Day", Comparisons))

KZFP_Inter <- KZFP_all %>%
  filter(grepl("^Sig_Sp_inter_", Comparisons))

KZFP_HvC <- KZFP_all %>%
  filter(grepl("^HvC_", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

ggplot(
  KZFP_HvC,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  theme_bw() +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkblue"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 timepoint"
    )
  ) +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "KZFP status",
    title = "Human vs. Chimpanzee expression of KZFPs"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

KZFP_Human_Traj <- KZFP_all %>%
  filter(grepl("^H_Day", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

KZFP_Human_Traj$Comparisons <- factor(
  KZFP_Human_Traj$Comparisons,
  levels = c(
    "H_Day02",
    "H_Day25",
    "H_Day515",
    "H_Day1530"
  )
)

ggplot(
  KZFP_Human_Traj,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkred"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 transition"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "KZFP status",
    title = "Human KZFP trajectories"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


KZFP_Chimp_Traj <- KZFP_all %>%
  filter(grepl("^C_Day", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

KZFP_Chimp_Traj$Comparisons <- factor(
  KZFP_Chimp_Traj$Comparisons,
  levels = c(
    "C_Day02",
    "C_Day25",
    "C_Day515",
    "C_Day1530"
  )
)

ggplot(
  KZFP_Chimp_Traj,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkred"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 transition"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "KZFP status",
    title = "Chimpanzee KZFP trajectories"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )




KZFP_Inter <- KZFP_all %>%
  filter(grepl("^Sig_Sp_inter_", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

KZFP_Inter$Comparisons <- factor(
  KZFP_Inter$Comparisons,
  levels = c(
    "Sig_Sp_inter_Traj02",
    "Sig_Sp_inter_Traj25",
    "Sig_Sp_inter_Traj515",
    "Sig_Sp_inter_Traj1530"
  )
)

ggplot(
  KZFP_Inter,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkgreen"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 interaction"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "KZFP status",
    title = "Species × trajectory interactions"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r TF_Species_FC_lineplot}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

TF_all <- TopTable %>%
  filter(
    Comparisons %in% comparison_order,
    Symbol %in% TF_list
  ) %>%
  mutate(
    Comparisons = factor(
      Comparisons,
      levels = comparison_order
    )
  )

library(dplyr)
library(ggplot2)

# Split the existing TF data into the three comparison types
TF_HvC <- TF_all %>%
  filter(grepl("^HvC_", Comparisons))

TF_Traj <- TF_all %>%
  filter(grepl("^[HC]_Day", Comparisons))

TF_Inter <- TF_all %>%
  filter(grepl("^Sig_Sp_inter_", Comparisons))

TF_HvC <- TF_all %>%
  filter(grepl("^HvC_", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

ggplot(
  TF_HvC,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  theme_bw() +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkblue"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 timepoint"
    )
  ) +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "TF status",
    title = "Human vs. Chimpanzee expression of TF"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

TF_Human_Traj <- TF_all %>%
  filter(grepl("^H_Day", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

TF_Human_Traj$Comparisons <- factor(
  TF_Human_Traj$Comparisons,
  levels = c(
    "H_Day02",
    "H_Day25",
    "H_Day515",
    "H_Day1530"
  )
)

ggplot(
  TF_Human_Traj,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkred"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 transition"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_cartesian(ylim = c(-10, 10)) +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "TF status",
    title = "Human TF trajectories"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


TF_Chimp_Traj <- TF_all %>%
  filter(grepl("^C_Day", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

TF_Chimp_Traj$Comparisons <- factor(
  TF_Chimp_Traj$Comparisons,
  levels = c(
    "C_Day02",
    "C_Day25",
    "C_Day515",
    "C_Day1530"
  )
)

ggplot(
  TF_Chimp_Traj,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkred"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 transition"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_cartesian(ylim = c(-10, 10)) +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "TF status",
    title = "Chimpanzee TF trajectories"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )




TF_Inter <- TF_all %>%
  filter(grepl("^Sig_Sp_inter_", Comparisons)) %>%
  group_by(Symbol) %>%
  mutate(
    DEG_any = any(!is.na(adj.P.Val) & adj.P.Val < 0.05)
  ) %>%
  ungroup()

TF_Inter$Comparisons <- factor(
  TF_Inter$Comparisons,
  levels = c(
    "Sig_Sp_inter_Traj02",
    "Sig_Sp_inter_Traj25",
    "Sig_Sp_inter_Traj515",
    "Sig_Sp_inter_Traj1530"
  )
)

ggplot(
  TF_Inter,
  aes(
    x = Comparisons,
    y = logFC,
    group = Symbol,
    color = DEG_any
  )
) +
  geom_line(alpha = 0.5) +
  geom_point(size = 1.5) +
  scale_color_manual(
    values = c(
      "FALSE" = "grey70",
      "TRUE" = "darkgreen"
    ),
    labels = c(
      "FALSE" = "Not DEG",
      "TRUE" = "DEG in ≥1 interaction"
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    x = "Comparison",
    y = "logFC",
    color = "TF status",
    title = "Species × trajectory interactions"
  ) +
  coord_cartesian(ylim = c(-10, 10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
#### Alluvial plot

```{r KZFP_Species_alluvial,fig.height=5,fig.width=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

comparisons <- unique(TopTable$Comparisons)

comparison_order <- c(
  "HvC_Day0",
  "HvC_Day2",
  "HvC_Day5",
  "HvC_Day15",
  "HvC_Day30",
  "H_Day02",
  "H_Day25",
  "H_Day515",
  "H_Day1530",
  "C_Day02",
  "C_Day25",
  "C_Day515",
  "C_Day1530",
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530"
)


KZFP_alluvial <- TopTable %>%
  filter(
    Symbol %in% KZFPs_list,
    Comparisons %in% comparison_order
  ) %>%
  mutate(
    Direction = case_when(
      adj.P.Val < 0.05 & logFC > 0 ~ "Up",
      adj.P.Val < 0.05 & logFC < 0 ~ "Down",
      TRUE ~ "Not significant"
    ),
    Comparison_Group = case_when(
      grepl("^HvC", Comparisons) ~ "Species",
      grepl("^H_Day", Comparisons) ~ "Human",
      grepl("^C_Day", Comparisons) ~ "Chimp",
      grepl("^Sig_Sp_inter", Comparisons) ~ "Interaction",
      TRUE ~ "Other"
    ),
    Comparisons = factor(
      Comparisons,
      levels = comparison_order
    )
  )

plot_KZFP_alluvial <- function(data, group_to_plot) {
  
  plot_data <- data %>%
    filter(Comparison_Group == group_to_plot)
  
  ggplot(
    plot_data,
    aes(
      x = Comparisons,
      stratum = Direction,
      alluvium = Symbol,
      y = 1,
      fill = Direction
    )
  ) +
    geom_alluvium(alpha = 0.5) +
    geom_stratum() +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      x = NULL,
      y = "Number of KZFPs",
      title = group_to_plot
    )
}



Species_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Species"
)

Human_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Human"
)

Chimp_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Chimp"
)

Interaction_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Interaction"
)

Species_plot
Human_plot
Chimp_plot
Interaction_plot



```

### Species

```{r KZFP_Species,fig.height=4, fig.width=7}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
TopTable_HvC_DEG_Fate <- readRDS("data/DGE/TopTable_HvC_DEG_Fate.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

# Add fate annotation to full TopTable
TopTable_Fate <- TopTable %>%
  filter(
    Comparisons %in% c(
      "HvC_Day0",
      "HvC_Day2",
      "HvC_Day5",
      "HvC_Day15",
      "HvC_Day30"
    )
  ) %>%
  left_join(
    TopTable_HvC_DEG_Fate %>%
      dplyr::select(Gene, Comparisons, Fate),
    by = c("Gene", "Comparisons")
  ) %>%
  mutate(
    Fate = ifelse(is.na(Fate), "Non-DEG", Fate)
  )

comparisons <- unique(TopTable_Fate$Comparisons)

Fate_KZFP_results <- Fate_KZFP_enrichment(
  comparisons = unique(TopTable_Fate$Comparisons),
  TopTable_Fate = TopTable_Fate,
  KZFPs_list = KZFPs_list
)


# FDR correction within each HvC comparison
Fate_KZFP_results <- Fate_KZFP_results %>%
  group_by(Comparison) %>%
  mutate(
    FDR = p.adjust(Fisher_p, method = "BH")
  ) %>%
  ungroup()


# Order developmental stages
comparison_order <- c(
  "HvC_Day0",
  "HvC_Day2",
  "HvC_Day5",
  "HvC_Day15",
  "HvC_Day30"
)


Fate_KZFP_results <- Fate_KZFP_results %>%
  mutate(
    Comparison = factor(
      Comparison,
      levels = comparison_order
    ),
    Significant = FDR < 0.05
  )

Fate_KZFP_results

plot_Fate_OR(Fate_KZFP_results)

```



```{r KZFP_Species_2}
KZFP_Toptable <- readRDS("data/TEs/KZFPs_Toptable.RDS")

KZFP_Toptable_HvC_Day0_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "HvC_Day0") %>% 
  filter(adj.P.Val < 0.05)
Day0_Sig_DEG <- KZFP_Toptable_HvC_Day0_DEG$Symbol

KZFP_Toptable_HvC_Day2_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "HvC_Day2") %>% 
  filter(adj.P.Val < 0.05) %>% 
  filter(logFC < 0)
Day2_Sig_DEG <- KZFP_Toptable_HvC_Day2_DEG$Symbol
Day2_Sig_DEG

KZFP_Toptable_HvC_Day5_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "HvC_Day5") %>% 
  filter(adj.P.Val < 0.05) %>% 
  filter(logFC < 0)
Day5_Sig_DEG <- KZFP_Toptable_HvC_Day5_DEG$Symbol

KZFP_Toptable_HvC_Day15_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "HvC_Day15") %>% 
  filter(adj.P.Val < 0.05) %>% 
  filter(logFC < 0)
Day15_Sig_DEG <- KZFP_Toptable_HvC_Day15_DEG$Symbol

KZFP_Toptable_HvC_Day30_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "HvC_Day30") %>% 
  filter(adj.P.Val < 0.05) %>% 
  filter(logFC < 0)
Day30_Sig_DEG <- KZFP_Toptable_HvC_Day30_DEG$Symbol

Day0_Sig_DEG_Unique <- setdiff(
  Day0_Sig_DEG,
  c(
    Day2_Sig_DEG,
    Day5_Sig_DEG,
    Day15_Sig_DEG,
    Day30_Sig_DEG
  )
)


  
```

#### Venn and Proportions

```{r KZFP_Species_Venn, fig.height=7,fig.width=7}
TopTable <- readRDS("data/TEs/KZFPs_Toptable.RDS")
RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

library(dplyr)
library(stringr)

timepoints <- c("Day0","Day2","Day5","Day15","Day30")

df <- TopTable %>%
  filter(str_detect(Comparisons, "^HvC_Day")) %>%
  mutate(
    Time = str_extract(Comparisons, "Day[0-9]+"),
    DEG = adj.P.Val < 0.05
  ) %>%
  filter(DEG)

genes_list <- list(
  Day0  = df %>% filter(Time == "Day0") %>% pull(Gene) %>% unique(),
  Day2  = df %>% filter(Time == "Day2") %>% pull(Gene) %>% unique(),
  Day5  = df %>% filter(Time == "Day5") %>% pull(Gene) %>% unique(),
  Day15 = df %>% filter(Time == "Day15") %>% pull(Gene) %>% unique(),
  Day30 = df %>% filter(Time == "Day30") %>% pull(Gene) %>% unique()
)

ggVennDiagram(genes_list, label = "count") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.text = element_text(size = 10),
    plot.title = element_text(face = "bold")
  )


df <- TopTable %>%
  filter(str_detect(Comparisons, "^HvC_Day")) %>%
  mutate(
    Stage = str_extract(Comparisons, "Day[0-9]+"),
    DEG = adj.P.Val < 0.05
  ) %>%
  filter(DEG)

genes <- list(
  Day0  = df %>% filter(Stage == "Day0") %>% pull(Gene) %>% unique(),
  Day2  = df %>% filter(Stage == "Day2") %>% pull(Gene) %>% unique(),
  Day5  = df %>% filter(Stage == "Day5") %>% pull(Gene) %>% unique(),
  Day15 = df %>% filter(Stage == "Day15") %>% pull(Gene) %>% unique(),
  Day30 = df %>% filter(Stage == "Day30") %>% pull(Gene) %>% unique()
)

days <- names(genes)

shared_all <- Reduce(intersect, genes)

get_fate <- function(day) {

  g <- genes[[day]]

  specific <- setdiff(
    g,
    unlist(genes[names(genes) != day])
  )

  shared_all_day <- intersect(
    g,
    shared_all
  )

  shared_partial <- setdiff(
    setdiff(g, specific),
    shared_all_day
  )

  data.frame(
    Day = day,
    Specific = length(specific),
    Shared_partial = length(shared_partial),
    Shared_all = length(shared_all_day)
  )
}

plot_df <- do.call(
  rbind,
  lapply(names(genes), get_fate)
)

library(tidyr)

plot_long <- plot_df %>%
  pivot_longer(
    cols = c(Specific, Shared_partial, Shared_all),
    names_to = "Category",
    values_to = "Count"
  )

plot_long$Day <- factor(
  plot_long$Day,
  levels = c("Day0","Day2","Day5","Day15","Day30")
)

library(ggplot2)

ggplot(plot_long, aes(x = Day, y = Count, fill = Category)) +
  geom_col() +
  scale_fill_manual(values = c(
    "Specific" = "#D55E00",
    "Shared_partial" = "#0072B2",
    "Shared_all" = "#009E73"
  )) +
  theme_bw() +
  labs(
    x = "Developmental stage",
    y = "Number of DEGs",
    fill = "",
    title = "Fate of HvC KZFP DEGs across development"
  )


library(dplyr)
library(tidyr)

plot_long_prop <- plot_df %>%
  pivot_longer(
    cols = c(Specific, Shared_partial, Shared_all),
    names_to = "Category",
    values_to = "Count"
  ) %>%
  group_by(Day) %>%
  mutate(
    Prop = Count / sum(Count)
  ) %>%
  ungroup()

plot_long_prop$Day <- factor(
  plot_long_prop$Day,
  levels = c("Day0","Day2","Day5","Day15","Day30")
)

library(ggplot2)

ggplot(plot_long_prop, aes(x = Day, y = Prop, fill = Category)) +
  geom_col() +
  scale_fill_manual(values = c(
    "Specific" = "#D55E00",
    "Shared_partial" = "#0072B2",
    "Shared_all" = "#009E73"
  )) +
  theme_bw() +
  labs(
    x = "Developmental stage",
    y = "Proportion of DEGs",
    fill = "",
    title = "Fate of HvC KZFP DEGs across development (normalized)"
  )

```
```{r RNASeq_DGE_Species_Venn_RLCOMP_G1_4SpTp_No4}
library(dplyr)
library(purrr)
library(stringr)

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs <- readRDS("data/TEs/KZFPs.RDS")
KZFP_list <- KZFPs$`Gene ID`

RNA_sets_KZFP <- RNA_sets

for (name in names(RNA_sets_KZFP)) {
  
  RNA_sets_KZFP[[name]]$DEGs <- RNA_sets_KZFP[[name]]$DEGs %>%
    filter(Gene %in% KZFP_list)
  
  RNA_sets_KZFP[[name]]$NonDEGs <- RNA_sets_KZFP[[name]]$NonDEGs %>%
    filter(Gene %in% KZFP_list)
}

hvcdays <- names(RNA_sets_KZFP)[str_detect(names(RNA_sets_KZFP), "^HvC_")]



prop_df <- map_dfr(hvcdays, function(traj) {
  
  x <- RNA_sets_KZFP[[traj]]
  
  tibble(
    trajectory = traj,
    DEG = length(x$DEGs$Gene),
    NonDEG = length(x$NonDEGs$Gene)
  )
})
library(tidyr)

prop_long <- prop_df %>%
  pivot_longer(cols = c(DEG, NonDEG),
               names_to = "type",
               values_to = "count") %>%
  group_by(trajectory) %>%
  mutate(prop = count / sum(count)) %>%
  ungroup()
prop_long$trajectory <- factor(
  prop_long$trajectory,
  levels = c("HvC_Day0", "HvC_Day2", "HvC_Day5", "HvC_Day15", "HvC_Day30"),
  labels = c("0", "2", "5", "15", "30")
)
library(ggplot2)

ggplot(prop_long, aes(x = trajectory, y = prop, fill = type)) +
  geom_col() +
  theme_bw() +
  labs(
    x = "Time",
    y = "Proportion of KZFPs",
    fill = "",
    title = "HvC comparisons: KZFP DEG vs KZFP NonDEG composition"
  )
```


### Trajectory

```{r KZFP_species_intersects}
KZFP_Toptable <- readRDS("data/TEs/KZFPs_Toptable.RDS")

KZFP_Toptable_sig <- KZFP_Toptable %>% 
  filter(adj.P.Val < 0.05)

# write.csv(KZFP_Toptable_sig, "KZFP_Toptable_sig.csv", row.names = FALSE)

comparisons <- unique(KZFP_Toptable_sig$Comparisons)

for (comparison in comparisons) {
  
  KZFP_Toptable_Higher <- KZFP_Toptable %>% 
    filter(Comparisons == comparison) %>% 
    filter(adj.P.Val < 0.05) %>% 
    filter(logFC > 0)
  
  KZFP_Toptable_Lower <- KZFP_Toptable %>% 
    filter(Comparisons == comparison) %>% 
    filter(adj.P.Val < 0.05) %>% 
    filter(logFC < 0)
  
  assign(
    paste0("KZFP_Toptable_", comparison, "_Higher"),
    KZFP_Toptable_Higher
  )
  
  assign(
    paste0("KZFP_Toptable_", comparison, "_Lower"),
    KZFP_Toptable_Lower
  )
}

Reduce(
  intersect,
  list(
    KZFP_Toptable_HvC_Day0_Higher$Symbol,
    KZFP_Toptable_HvC_Day2_Higher$Symbol,
    KZFP_Toptable_HvC_Day5_Higher$Symbol,
    KZFP_Toptable_HvC_Day15_Higher$Symbol,
    KZFP_Toptable_HvC_Day30_Higher$Symbol
  )
)

Reduce(
  intersect,
  list(
    KZFP_Toptable_HvC_Day0_Lower$Symbol,
    KZFP_Toptable_HvC_Day2_Lower$Symbol,
    KZFP_Toptable_HvC_Day5_Lower$Symbol,
    KZFP_Toptable_HvC_Day15_Lower$Symbol,
    KZFP_Toptable_HvC_Day30_Lower$Symbol
  )
)
```
#### Proportions

```{r RNASeq_DGE_Trajectory_Proportions_RLCOMP_G1_4SpTp_No4_4, fig.height=7,fig.width=7}
TopTable <- readRDS("data/DGE/TopTable_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")



traj_levels <- c("H_Day02","H_Day25","H_Day515","H_Day1530",
                 "C_Day02","C_Day25","C_Day515","C_Day1530")

df <- TopTable %>%
  dplyr::filter(Comparisons %in% traj_levels) %>% 
  dplyr::filter(Gene %in% KZFP_list)

df <- df %>%
  dplyr::mutate(
    Species = ifelse(grepl("^H", Comparisons), "Human", "Chimp"),
    Stage = sub(".*_(Day[0-9]+)", "\\1", Comparisons),
    DEG = adj.P.Val < 0.05
  )

df_deg <- df %>%
  dplyr::filter(DEG)

library(dplyr)

df_deg2 <- df_deg %>%
  group_by(Gene, Stage, Species) %>%
  summarise(.groups = "drop")

df_overlap <- df_deg2 %>%
  tidyr::pivot_wider(
    names_from = Species,
    values_from = Gene,
    values_fn = length,
    values_fill = 0
  )

df_flags <- df_deg2 %>%
  mutate(flag = TRUE) %>%
  tidyr::pivot_wider(
    names_from = Species,
    values_from = flag,
    values_fill = FALSE
  ) %>%
  dplyr::mutate(
    category = case_when(
      Human & Chimp ~ "Shared",
      Human ~ "Human only",
      Chimp ~ "Chimp only"
    )
  )

plot_df <- df_flags %>%
  group_by(Stage, category) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Stage) %>%
  mutate(prop = n / sum(n))

plot_df$Stage <- factor(
  plot_df$Stage,
  levels = c("Day02", "Day25", "Day515", "Day1530")
)

library(ggpattern)

ggplot(plot_df, aes(x = Stage, y = prop, fill = category)) +

  geom_col_pattern(
    aes(pattern = category),
    colour = "black",
    pattern_colour = "black",
    pattern_fill = "white",
    pattern_density = 0.4,
    pattern_spacing = 0.02
  ) +

  scale_fill_manual(values = c(
    "Shared" = "grey70",
    "Human only" = "black",
    "Chimp only" = "white"
  )) +

  scale_pattern_manual(values = c(
    "Shared" = "stripe",
    "Human only" = "none",
    "Chimp only" = "none"
  )) +

  theme_minimal(base_size = 14) +
  labs(
    x = "Developmental Stage",
    y = "Proportion of KZFP DEGs",
    fill = "",
    pattern = ""
  )

ggplot(plot_df, aes(x = Stage, y = n, fill = category)) +

  geom_col_pattern(
    aes(pattern = category),
    colour = "black",
    pattern_colour = "black",
    pattern_fill = "white",
    pattern_density = 0.4,
    pattern_spacing = 0.02
  ) +

  scale_fill_manual(values = c(
    "Shared" = "grey70",
    "Human only" = "black",
    "Chimp only" = "white"
  )) +

  scale_pattern_manual(values = c(
    "Shared" = "stripe",
    "Human only" = "none",
    "Chimp only" = "none"
  )) +

  theme_minimal(base_size = 14) +
  labs(
    x = "Developmental Stage",
    y = "Number of KZFP DEGs",
    fill = "",
    pattern = ""
  )
```

### Interaction

```{r KZFP_Interactions,fig.height=4, fig.width=7}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
TopTable_HvC_DEG_Fate <- readRDS("data/DGE/TopTable_HvC_DEG_Fate.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

# Add fate annotation to full TopTable
TopTable_Interaction <- TopTable %>%
  filter(
    Comparisons %in% c(
      "Sig_Sp_inter_Traj02",
      "Sig_Sp_inter_Traj25",
      "Sig_Sp_inter_Traj515",
      "Sig_Sp_inter_Traj1530"
    ))

```

#### FC and absFC

```{r KZFP_Interactions_FC}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

comparisons <- unique(TopTable$Comparisons)

comparison_order <- c(
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530"
)

KZFP_interaction <- TopTable %>%
  filter(
    Comparisons %in% comparison_order,
    Symbol %in% KZFPs_list
  ) %>%
  mutate(
    Comparisons = factor(
      Comparisons,
      levels = comparison_order
    )
  )

# saveRDS(KZFP_all,"data/TEs/KZFPs_Toptable.RDS")

KZFP_interaction$absFC <- abs(KZFP_interaction$logFC)

KZFP_interaction %>%
  distinct(Symbol, Comparisons) %>%
  count(Symbol, name = "n_comparisons") %>%
  count(n_comparisons)

pairwise.wilcox.test(
  KZFP_interaction$logFC,
  KZFP_interaction$Comparisons,
  paired = TRUE,
  p.adjust.method = "BH"
)

ggplot(KZFP_interaction, aes(Comparisons, logFC)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

friedman.test(
  logFC ~ Comparisons | Symbol,
  data = KZFP_interaction
)

ggplot(KZFP_interaction, aes(Comparisons, absFC)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

friedman.test(
  absFC ~ Comparisons | Symbol,
  data = KZFP_interaction
)

pairwise.wilcox.test(
  KZFP_interaction$absFC,
  KZFP_interaction$Comparisons,
  paired = TRUE,
  p.adjust.method = "BH"
)
```


```{r KZFP_Interactions_directions}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

comparisons <- c(
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530"
)

KZFP_FC <- TopTable %>%
  filter(
    Comparisons %in% comparisons,
    Symbol %in% KZFPs_list
  ) %>%
  dplyr::select(Comparisons, Symbol, logFC, adj.P.Val)

KZFP_DEG <- TopTable %>%
  filter(
    Comparisons %in% comparisons,
    Symbol %in% KZFPs_list,
    adj.P.Val < 0.05
  ) %>%
  dplyr::select(Comparisons, Symbol, logFC, adj.P.Val)

KZFP_DEG %>%
  arrange(Comparisons, desc(logFC))

KZFP_DEG_direction <- KZFP_DEG %>%
  mutate(
    Direction = case_when(
      # Species x trajectory interactions
      grepl("^Sig_Sp_inter", Comparisons) & logFC > 0 ~ "Human-Chimp difference increased",
      grepl("^Sig_Sp_inter", Comparisons) & logFC < 0 ~ "Human-Chimp difference decreased",
      
      TRUE ~ "Unknown"
    )
  ) %>%
  dplyr::count(Comparisons, Direction)

KZFP_DEG_direction
```

#### Alluvial plot

```{r KZFP_Interactions_alluvial,fig.height=5,fig.width=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

comparisons <- unique(TopTable$Comparisons)

comparison_order <- c(
  "HvC_Day0",
  "HvC_Day2",
  "HvC_Day5",
  "HvC_Day15",
  "HvC_Day30",
  "H_Day02",
  "H_Day25",
  "H_Day515",
  "H_Day1530",
  "C_Day02",
  "C_Day25",
  "C_Day515",
  "C_Day1530",
  "Sig_Sp_inter_Traj02",
  "Sig_Sp_inter_Traj25",
  "Sig_Sp_inter_Traj515",
  "Sig_Sp_inter_Traj1530"
)


KZFP_alluvial <- TopTable %>%
  filter(
    Symbol %in% KZFPs_list,
    Comparisons %in% comparison_order
  ) %>%
  mutate(
    Direction = case_when(
      adj.P.Val < 0.05 & logFC > 0 ~ "Up",
      adj.P.Val < 0.05 & logFC < 0 ~ "Down",
      TRUE ~ "Not significant"
    ),
    Comparison_Group = case_when(
      grepl("^HvC", Comparisons) ~ "Species",
      grepl("^H_Day", Comparisons) ~ "Human",
      grepl("^C_Day", Comparisons) ~ "Chimp",
      grepl("^Sig_Sp_inter", Comparisons) ~ "Interaction",
      TRUE ~ "Other"
    ),
    Comparisons = factor(
      Comparisons,
      levels = comparison_order
    )
  )

plot_KZFP_alluvial <- function(data, group_to_plot) {
  
  plot_data <- data %>%
    filter(Comparison_Group == group_to_plot)
  
  ggplot(
    plot_data,
    aes(
      x = Comparisons,
      stratum = Direction,
      alluvium = Symbol,
      y = 1,
      fill = Direction
    )
  ) +
    geom_alluvium(alpha = 0.5) +
    geom_stratum() +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      x = NULL,
      y = "Number of KZFPs",
      title = group_to_plot
    )
}



Species_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Species"
)

Human_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Human"
)

Chimp_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Chimp"
)

Interaction_plot <- plot_KZFP_alluvial(
  KZFP_alluvial,
  "Interaction"
)

Species_plot
Human_plot
Chimp_plot
Interaction_plot



```

#### Venn and Proportions

```{r KZFP_Interactions_Venn, fig.height=7,fig.width=7}
TopTable <- readRDS("data/TEs/KZFPs_Toptable.RDS")
RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

library(dplyr)
library(stringr)

timepoints <- c("Day0","Day2","Day5","Day15","Day30")

df <- TopTable %>%
  filter(str_detect(Comparisons, "^HvC_Day")) %>%
  mutate(
    Time = str_extract(Comparisons, "Day[0-9]+"),
    DEG = adj.P.Val < 0.05
  ) %>%
  filter(DEG)

genes_list <- list(
  Day0  = df %>% filter(Time == "Day0") %>% pull(Gene) %>% unique(),
  Day2  = df %>% filter(Time == "Day2") %>% pull(Gene) %>% unique(),
  Day5  = df %>% filter(Time == "Day5") %>% pull(Gene) %>% unique(),
  Day15 = df %>% filter(Time == "Day15") %>% pull(Gene) %>% unique(),
  Day30 = df %>% filter(Time == "Day30") %>% pull(Gene) %>% unique()
)

ggVennDiagram(genes_list, label = "count") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.text = element_text(size = 10),
    plot.title = element_text(face = "bold")
  )


df <- TopTable %>%
  filter(str_detect(Comparisons, "^HvC_Day")) %>%
  mutate(
    Stage = str_extract(Comparisons, "Day[0-9]+"),
    DEG = adj.P.Val < 0.05
  ) %>%
  filter(DEG)

genes <- list(
  Day0  = df %>% filter(Stage == "Day0") %>% pull(Gene) %>% unique(),
  Day2  = df %>% filter(Stage == "Day2") %>% pull(Gene) %>% unique(),
  Day5  = df %>% filter(Stage == "Day5") %>% pull(Gene) %>% unique(),
  Day15 = df %>% filter(Stage == "Day15") %>% pull(Gene) %>% unique(),
  Day30 = df %>% filter(Stage == "Day30") %>% pull(Gene) %>% unique()
)

days <- names(genes)

shared_all <- Reduce(intersect, genes)

get_fate <- function(day) {

  g <- genes[[day]]

  specific <- setdiff(
    g,
    unlist(genes[names(genes) != day])
  )

  shared_all_day <- intersect(
    g,
    shared_all
  )

  shared_partial <- setdiff(
    setdiff(g, specific),
    shared_all_day
  )

  data.frame(
    Day = day,
    Specific = length(specific),
    Shared_partial = length(shared_partial),
    Shared_all = length(shared_all_day)
  )
}

plot_df <- do.call(
  rbind,
  lapply(names(genes), get_fate)
)

library(tidyr)

plot_long <- plot_df %>%
  pivot_longer(
    cols = c(Specific, Shared_partial, Shared_all),
    names_to = "Category",
    values_to = "Count"
  )

plot_long$Day <- factor(
  plot_long$Day,
  levels = c("Day0","Day2","Day5","Day15","Day30")
)

library(ggplot2)

ggplot(plot_long, aes(x = Day, y = Count, fill = Category)) +
  geom_col() +
  scale_fill_manual(values = c(
    "Specific" = "#D55E00",
    "Shared_partial" = "#0072B2",
    "Shared_all" = "#009E73"
  )) +
  theme_bw() +
  labs(
    x = "Developmental stage",
    y = "Number of DEGs",
    fill = "",
    title = "Fate of HvC KZFP DEGs across development"
  )


library(dplyr)
library(tidyr)

plot_long_prop <- plot_df %>%
  pivot_longer(
    cols = c(Specific, Shared_partial, Shared_all),
    names_to = "Category",
    values_to = "Count"
  ) %>%
  group_by(Day) %>%
  mutate(
    Prop = Count / sum(Count)
  ) %>%
  ungroup()

plot_long_prop$Day <- factor(
  plot_long_prop$Day,
  levels = c("Day0","Day2","Day5","Day15","Day30")
)

library(ggplot2)

ggplot(plot_long_prop, aes(x = Day, y = Prop, fill = Category)) +
  geom_col() +
  scale_fill_manual(values = c(
    "Specific" = "#D55E00",
    "Shared_partial" = "#0072B2",
    "Shared_all" = "#009E73"
  )) +
  theme_bw() +
  labs(
    x = "Developmental stage",
    y = "Proportion of DEGs",
    fill = "",
    title = "Fate of HvC KZFP DEGs across development (normalized)"
  )

```

```{r KZFP_Interactions_lists}
KZFP_Toptable <- readRDS("data/TEs/KZFPs_Toptable.RDS")

KZFP_Toptable_Sig_Sp_inter_Traj02_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "Sig_Sp_inter_Traj02") %>% 
  filter(adj.P.Val < 0.05)
Sig_Sp_inter_Traj02_Sig_DEG <-KZFP_Toptable_Sig_Sp_inter_Traj02_DEG$Symbol

KZFP_Toptable_Sig_Sp_inter_Traj25_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "Sig_Sp_inter_Traj25") %>% 
  filter(adj.P.Val < 0.05)
Sig_Sp_inter_Traj25_Sig_DEG <-KZFP_Toptable_Sig_Sp_inter_Traj25_DEG$Symbol

KZFP_Toptable_Sig_Sp_inter_Traj515_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "Sig_Sp_inter_Traj515") %>% 
  filter(adj.P.Val < 0.05)
Sig_Sp_inter_Traj515_Sig_DEG <-KZFP_Toptable_Sig_Sp_inter_Traj515_DEG$Symbol

KZFP_Toptable_Sig_Sp_inter_Traj1530_DEG <- KZFP_Toptable %>% 
  filter(Comparisons == "Sig_Sp_inter_Traj1530") %>% 
  filter(adj.P.Val < 0.05)
Sig_Sp_inter_Traj1530_Sig_DEG <-KZFP_Toptable_Sig_Sp_inter_Traj1530_DEG$Symbol


Sig_Sp_inter_Traj02_Sig_DEG_specific <- setdiff(
  Sig_Sp_inter_Traj02_Sig_DEG,
  c(
    Sig_Sp_inter_Traj25_Sig_DEG,
    Sig_Sp_inter_Traj515_Sig_DEG,
    Sig_Sp_inter_Traj1530_Sig_DEG
  )
)
Sig_Sp_inter_Traj02_Sig_DEG_specific


Sig_Sp_inter_Traj25_Sig_DEG_specific <- setdiff(
  Sig_Sp_inter_Traj25_Sig_DEG,
  c(
    Sig_Sp_inter_Traj02_Sig_DEG,
    Sig_Sp_inter_Traj515_Sig_DEG,
    Sig_Sp_inter_Traj1530_Sig_DEG
  )
)
Sig_Sp_inter_Traj25_Sig_DEG_specific

Sig_Sp_inter_Traj515_Sig_DEG_specific <- setdiff(
  Sig_Sp_inter_Traj515_Sig_DEG,
  c(
    Sig_Sp_inter_Traj02_Sig_DEG,
    Sig_Sp_inter_Traj25_Sig_DEG,
    Sig_Sp_inter_Traj1530_Sig_DEG
  )
)
Sig_Sp_inter_Traj515_Sig_DEG_specific
  
Sig_Sp_inter_Traj1530_Sig_DEG_specific <- setdiff(
  Sig_Sp_inter_Traj1530_Sig_DEG,
  c(
    Sig_Sp_inter_Traj02_Sig_DEG,
    Sig_Sp_inter_Traj25_Sig_DEG,
    Sig_Sp_inter_Traj515_Sig_DEG
  )
)
Sig_Sp_inter_Traj1530_Sig_DEG_specific

Sig_Sp_inter_Sig_DEG_nonspecific <- intersect(
  Sig_Sp_inter_Traj02_Sig_DEG,
    c(Sig_Sp_inter_Traj25_Sig_DEG,
    Sig_Sp_inter_Traj515_Sig_DEG,
    Sig_Sp_inter_Traj1530_Sig_DEG
  ))
Sig_Sp_inter_Sig_DEG_nonspecific
```


### CorMotif Species

#### Data Adding
```{r KZFP_Interactions_Cormotif,fig.height=4, fig.width=7}
cluster_annotation <- readRDS("data/Cormotif/Species/SpeciesClusterAnnotation_logcpm_G1_SpTp_TMM_No4_NoRep.RDS")


TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TopTable$Cluster <- unname(cluster_annotation[rownames(TopTable)])

TopTable$Cluster <- ifelse(
  is.na(TopTable$Cluster),
  "Unassigned",
  TopTable$Cluster
)

# saveRDS(TopTable,"data/DGE/TopTable_Symbols_Clusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
```

####Enrichment Unassigned 
```{r KZFP_Interactions_Cormotif_enrichment_Unassigned,fig.height=4, fig.width=4}

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

TopTable <- readRDS(
  "data/DGE/TopTable_Symbols_Clusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS"
)

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0")

TopTable_OneSet_Unassigned <- TopTable_OneSet %>% 
  filter(Cluster == "Unassigned")
length(unique(TopTable_OneSet$Gene))

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")


clusters <- paste0("cluster", 1:5)

KZFP_cluster_results <- KZFP_cluster_enrichment_Unassigned(
  clusters,
  TopTable_OneSet,
  KZFPs_list
)

KZFP_cluster_results$FDR <- p.adjust(
  KZFP_cluster_results$Fisher_p,
  method = "BH"
)

 KZFP_cluster_results
 
 plot_OR_cluster(
  KZFP_cluster_results,
  paste0("cluster", 1:5)
)
 
```

####Enrichment No Response
```{r KZFP_Interactions_Cormotif_enrichment_NoRepsonse,fig.height=4, fig.width=4}

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

TopTable <- readRDS(
  "data/DGE/TopTable_Symbols_Clusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS"
)

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0") %>% 
  filter(Cluster != "Unassigned")

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")


clusters <- paste0("cluster", 2:5)

KZFP_cluster_results <- cluster_enrichment_NoResponse(
  clusters,
  TopTable_OneSet,
  KZFPs_list
)

KZFP_cluster_results$FDR <- p.adjust(
  KZFP_cluster_results$Fisher_p,
  method = "BH"
)

 KZFP_cluster_results
 
  plot_OR_cluster(
  KZFP_cluster_results,
  paste0("cluster", 2:5)
)
 
```

#### TF Enrichment No Response
```{r TF_Interactions_Cormotif_enrichment_NoResponse,fig.height=4, fig.width=4}

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

TopTable <- readRDS(
  "data/DGE/TopTable_Symbols_Clusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS"
)

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0") %>% 
  filter(Cluster != "Unassigned")


clusters <- paste0("cluster", 2:5)

TF_cluster_results <- cluster_enrichment_NoResponse(
  clusters,
  TopTable_OneSet,
  TF_list
)

  TF_cluster_results$FDR <- p.adjust(
  TF_cluster_results$Fisher_p,
  method = "BH"
)

 TF_cluster_results
 
  plot_OR_cluster(
  TF_cluster_results,
  paste0("cluster", 2:5)
)
 
```

### CorMotif Trajectory

#### Data Adding
```{r KZFP_Cormotif_Trajectory,fig.height=4, fig.width=7}
cluster_annotation <- readRDS("data/Cormotif/Trajectory/TrajectoryClusterAnnotation_logcpm_G1_SpTp_TMM_No4_NoRep.RDS")


TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TopTable$Cluster <- unname(cluster_annotation[rownames(TopTable)])

TopTable$Cluster <- ifelse(
  is.na(TopTable$Cluster),
  "Unassigned",
  TopTable$Cluster
)

# saveRDS(TopTable,"data/DGE/TopTable_Symbols_TrajectoryClusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

```

####Enrichment Unassigned 
```{r KZFP_Cormotif_Trajectory_enrichment_Unassigned,fig.height=4, fig.width=4}

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

TopTable <- readRDS("data/DGE/TopTable_Symbols_TrajectoryClusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0")

TopTable_OneSet_Unassigned <- TopTable_OneSet %>% 
  filter(Cluster == "Unassigned")
length(unique(TopTable_OneSet$Gene))

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")


clusters <- paste0("cluster", 1:12)

KZFP_cluster_results <- KZFP_cluster_enrichment_Unassigned(
  clusters,
  TopTable_OneSet,
  KZFPs_list
)

KZFP_cluster_results$FDR <- p.adjust(
  KZFP_cluster_results$Fisher_p,
  method = "BH"
)

 KZFP_cluster_results
 
 plot_OR_cluster(
  KZFP_cluster_results,
  paste0("cluster", 1:12)
)
 
```

####Enrichment No Response
```{r KZFP_Cormotif_Trajectory_enrichment_NoRepsonse,fig.height=4, fig.width=4}

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")

TopTable <-readRDS("data/DGE/TopTable_Symbols_TrajectoryClusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0")

RNA_sets <- readRDS("data/DGE/RNA_sets_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs_list <- readRDS("data/TEs/KZFPs_list.RDS")


clusters <- paste0("cluster", 2:12)

KZFP_cluster_results <- cluster_enrichment_NoResponse(
  clusters,
  TopTable_OneSet,
  KZFPs_list
)

KZFP_cluster_results$FDR <- p.adjust(
  KZFP_cluster_results$Fisher_p,
  method = "BH"
)

 KZFP_cluster_results
 
  plot_OR_cluster(
  KZFP_cluster_results,
  paste0("cluster", 2:12)
)
 
```

#### TF Enrichment No Response
```{r TF_Cormotif_Trajectory_enrichment_NoResponse,fig.height=4, fig.width=4}

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

TopTable <-readRDS("data/DGE/TopTable_Symbols_TrajectoryClusters__Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TopTable_OneSet <- TopTable %>% 
  filter(Comparisons == "HvC_Day0") %>% 
  filter(Cluster != "Unassigned")


clusters <- paste0("cluster", 2:12)

TF_cluster_results <- cluster_enrichment_NoResponse(
  clusters,
  TopTable_OneSet,
  TF_list
)

  TF_cluster_results$FDR <- p.adjust(
  TF_cluster_results$Fisher_p,
  method = "BH"
)

 TF_cluster_results
 
  plot_OR_cluster(
  TF_cluster_results,
  paste0("cluster", 2:12)
)
 
```




```{r Updating Website, eval = FALSE, fig.height=10, fig.width=10}

# git -> commit all changes
# git -> push
wflow_publish("analysis/KZFP.Rmd")

```
