---
title: "RNA_CorrelationPlots"
output:
  html_document:
    toc: true
    toc_float: true
date: "2026-03-03"
---

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

# Correlation Plotting

## Setting up the environment

### Loading libraries
```{r library loading, include=FALSE, warning = FALSE, message=FALSE}

#loading libraries
library("tidyverse")
library("readxl")
library("readr")
library("ggrepel")
library("ggfortify")
library("patchwork")
library("eulerr")
library("org.Hs.eg.db")
library("AnnotationDbi")
library("ggVennDiagram")
library("ggrastr")
library("edgeR")
library("limma")
library("ggsignif")
library("dplyr")
library("purrr")
library("ggplot2")
library("tidyr")
# install.packages(c("knitr", "BiocManager"))
# BiocManager::install("org.Hs.eg.db")



#Loading R Objects Other WorkflowR

```

### Defining functions

```{r Functions_1, include=FALSE, warning = FALSE, message=FALSE}
# ---------------------------------------------------------
    # Setting up function variables
    # ---------------------------------------------------------

plot_species_comparison <- function(
    TopTable,
    comps,
    genes_to_include = NULL,
    genes_label = NULL,
    x_limits = c(-15, 15),
    y_limits = c(-15, 15)
) {

    # ---------------------------------------------------------
    # Actual plotting function
    # ---------------------------------------------------------
  make_plot <- function(comp) {

    H_comp <- paste0("H_", comp)
    C_comp <- paste0("C_", comp)
  
    # Prepare data
    df_wide <- TopTable %>%
      filter(Comparisons %in% c(H_comp, C_comp)) %>%
      dplyr::select(Gene, Symbol, Comparisons, logFC) %>%
    group_by(Gene, Symbol, Comparisons) %>%
    summarize(
      logFC = mean(logFC),
      .groups = "drop"
    ) %>%
    pivot_wider(
      names_from = Comparisons,
      values_from = logFC
    ) %>%
    filter(
      !is.na(.data[[H_comp]]),
      !is.na(.data[[C_comp]])
  )

    # ---------------------------------------------------------
    # Labeling for the purpose or color coding dots on plot
    # ---------------------------------------------------------

    if (!is.null(genes_to_include)) {

      df_wide <- df_wide %>%
        filter(Symbol %in% genes_to_include)

    }

    # Linear regression
    fit <- lm(
      df_wide[[C_comp]] ~ df_wide[[H_comp]]
    )

    r2 <- summary(fit)$r.squared
    pval <- summary(fit)$coefficients[2, "Pr(>|t|)"]

    # DEGs
    genes_of_interest_human <- TopTable %>%
  filter(
    Comparisons == H_comp,
    adj.P.Val < 0.05
  ) %>%
  pull(Symbol) %>%
  unique()

genes_of_interest_chimp <- TopTable %>%
  filter(
    Comparisons == C_comp,
    adj.P.Val < 0.05
  ) %>%
  pull(Symbol) %>%
  unique()

df_wide$highlight <- dplyr::case_when(

  df_wide$Symbol %in% genes_of_interest_human &
    df_wide$Symbol %in% genes_of_interest_chimp ~ "T-DEG Both",

  df_wide$Symbol %in% genes_of_interest_human ~ "T-DEG Human Only",

  df_wide$Symbol %in% genes_of_interest_chimp ~ "T-DEG Chimp Only",

  TRUE ~ "Other"
)

    # ---------------------------------------------------------
    # plotting
    # --------------------------------------------------------- 

    p <- ggplot(
      df_wide,
      aes(
        x = .data[[H_comp]],
        y = .data[[C_comp]],
        color = highlight
      )
    ) +

      geom_point(
        size = 3,
        alpha = 0.8
      ) +

      scale_color_manual(
        values = c(
          "T-DEG Human Only" = "red",
          "T-DEG Chimp Only" = "blue",
          "T-DEG Both" = "purple",
          "Other" = "grey70"
        )
      ) +

      geom_smooth(
        method = "lm",
        se = FALSE,
        linetype = "dashed",
        color = "black"
      ) +

      geom_hline(
        yintercept = 0,
        linetype = "dashed"
      ) +

      geom_vline(
        xintercept = 0,
        linetype = "dashed"
      ) +

      annotate(
        "text",
        x = x_limits[1],
        y = y_limits[2],
        label = paste0(
          "R² = ", round(r2, 2),
          "\np = ", format(
            pval,
            scientific = TRUE,
            digits = 3
          )
        ),
        hjust = 0,
        vjust = 1,
        size = 5
      ) +

      coord_cartesian(
        xlim = x_limits,
        ylim = y_limits
      ) +

      theme_bw() +

      labs(
        x = H_comp,
        y = C_comp,
        title = paste0(H_comp, " vs ", C_comp," --", genes_label),
        color = "Gene Set"
      )

    return(list(
  plot = p,
  r2 = r2,
  pval = pval
))
  }

  # Generate one plot for each comparison
  plots <- lapply(comps, make_plot)

  # Name plots
  names(plots) <- comps

  return(plots)
}
```


```{r Functions_2, include=FALSE, warning = FALSE, message=FALSE}
## Finding "matched" DEGs

    # ---------------------------------------------------------
    # Get genes significant in each comparison
    # --------------------------------------------------------- 

get_matched_degs <- function(
    TopTable,
    transitions,
    genes_to_include = NULL,
    FDR_cutoff = 0.05
) {

  matched_genes <- lapply(
    transitions,
    function(comp) {

      H_genes <- TopTable %>%
        filter(
          Comparisons == paste0("H_", comp),
          adj.P.Val < FDR_cutoff
        ) %>%
        pull(Symbol) %>%
        unique()

      C_genes <- TopTable %>%
        filter(
          Comparisons == paste0("C_", comp),
          adj.P.Val < FDR_cutoff
        ) %>%
        pull(Symbol) %>%
        unique()

      # Same gene is DEG in both species for this transition
      intersect(H_genes, C_genes)
    }
  )

  # Gene is retained if matched in at least one transition
  matched_genes <- unique(unlist(matched_genes))

  # Optional restriction to TFs, KZFPs, etc.
  if (!is.null(genes_to_include)) {
    matched_genes <- intersect(
      matched_genes,
      genes_to_include
    )
  }

  matched_genes
}
```



```{r Functions_3, include=FALSE, warning = FALSE, message=FALSE}


plot_R2_trajectories <- function(
    R2_table,
    colors
) {

  R2_table <- R2_table %>%
    mutate(
      Group = case_when(
        Gene_Set == "All Expressed Genes" ~ "All Expressed",
        Gene_Set == "All Trajectory DEGs" ~ "All DEGs",
        Gene_Set == "Species-Shared Trajectory DEGs" ~ "Shared DEGs",
        Gene_Set == "Species-NonShared Trajectory DEGs" ~ "Species-Specific DEGs",
        Gene_Set == "Trajectory Non-DEGs" ~ "Non-DEGs"
      ),
      
      Transition = case_when(
        Transition == "Day02" ~ "Day0-2",
        Transition == "Day24" ~ "Day2-4",
        Transition == "Day45" ~ "Day4-5",
        Transition == "Day515" ~ "Day5-15",
        Transition == "Day1530" ~ "Day15-30"
      )
    )
  # Set factor order
  R2_table$Transition <- factor(
    R2_table$Transition,
    levels = c(
      "Day0-2",
      "Day2-4",
      "Day4-5",
      "Day5-15",
      "Day15-30"
    )
  )

  R2_table$Group <- factor(
    R2_table$Group,
    levels = c(
      "All Expressed",
      "All DEGs",
      "Shared DEGs",
      "Species-Specific DEGs",
      "Non-DEGs"
    )
  )

  # Plot
  ggplot(
    R2_table,
    aes(
      x = Transition,
      y = R2,
      group = Group,
      linetype = Group
    )
  ) +

    annotate(
      "rect",
      xmin = 1.5,
      xmax = 3.5,
      ymin = -Inf,
      ymax = Inf,
      alpha = 0.1,
      fill = "grey70"
    ) +

    geom_line(
      aes(
        linewidth = Group,
        color = Group
      )
    ) +

 # Significant points
geom_point(
  data = R2_table %>% filter(P_value < 0.05),
  aes(color = Group),
  shape = 16,
  size = 3
) +

# Non-significant points
geom_point(
  data = R2_table %>% filter(P_value >= 0.05),
  aes(color = Group),
  shape = 4,
  size = 3,
  stroke = 1.2
) +
    # geom_text(
    #   aes(label = round(R2, 2)),
    #   vjust = -1,
    #   size = 3,
    #   show.legend = FALSE
    # ) +

    scale_linetype_manual(
      values = c(
        "All Expressed" = "solid",
        "All DEGs" = "solid",
        "Shared DEGs" = "longdash",
        "Species-Specific DEGs" = "dotdash",
        "Non-DEGs" = "dotted"
      )
    ) +

    scale_color_manual(
      values = colors
    ) +

    scale_linewidth_manual(
      values = c(
        "All Expressed" = 1.8,
        "All DEGs" = 1.8,
        "Shared DEGs" = 1.2,
        "Species-Specific DEGs" = 1.2,
        "Non-DEGs" = 1.2
      )
    ) +

    labs(
      title = "Inter-species correlation peaks at mid-development",
      x = "Differentiation Transition",
      y = expression(R^2),
      linetype = "Dataset",
      linewidth = "Dataset",
      color = "Dataset"
    ) +

    coord_cartesian(
      ylim = c(0, 1)
    ) +

    theme_minimal(
      base_size = 14
    ) +
    
    guides(
  color = "none",
  linetype = guide_legend(title = "Line Type"),
  linewidth = "none"
) +

    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(
        angle = 0,
        hjust = 0.5
      ),
      plot.title = element_text(face = "bold")
    )
}
```


```{r Functions_4, include=FALSE, warning = FALSE, message=FALSE}


plot_R2_trajectories_No4 <- function(
    R2_table,
    colors
) {

  R2_table <- R2_table %>%
    mutate(
      Group = case_when(
        Gene_Set == "All Expressed Genes" ~ "All Expressed",
        Gene_Set == "All Trajectory DEGs" ~ "All DEGs",
        Gene_Set == "Species-Shared Trajectory DEGs" ~ "Shared DEGs",
        Gene_Set == "Species-NonShared Trajectory DEGs" ~ "Species-Specific DEGs",
        Gene_Set == "Trajectory Non-DEGs" ~ "Non-DEGs"
      ),
      
      Transition = case_when(
        Transition == "Day02" ~ "Day0-2",
        Transition == "Day25" ~ "Day2-5",
        Transition == "Day515" ~ "Day5-15",
        Transition == "Day1530" ~ "Day15-30"
      )
    )
  # Set factor order
  R2_table$Transition <- factor(
    R2_table$Transition,
    levels = c(
      "Day0-2",
      "Day2-5",
      "Day5-15",
      "Day15-30"
    )
  )

  R2_table$Group <- factor(
    R2_table$Group,
    levels = c(
      "All Expressed",
      "All DEGs",
      "Shared DEGs",
      "Species-Specific DEGs",
      "Non-DEGs"
    )
  )

  # Plot
  ggplot(
    R2_table,
    aes(
      x = Transition,
      y = R2,
      group = Group,
      linetype = Group
    )
  ) +

    annotate(
      "rect",
      xmin = 1.5,
      xmax = 3.5,
      ymin = -Inf,
      ymax = Inf,
      alpha = 0.1,
      fill = "grey70"
    ) +

    geom_line(
      aes(
        linewidth = Group,
        color = Group
      )
    ) +

# Significant points
geom_point(
  data = R2_table %>% filter(P_value < 0.05),
  aes(color = Group),
  shape = 16,
  size = 3
) +

# Non-significant points
geom_point(
  data = R2_table %>% filter(P_value >= 0.05),
  aes(color = Group),
  shape = 4,
  size = 3,
  stroke = 1.2
) +

    # geom_text(
    #   aes(label = round(R2, 2)),
    #   vjust = -1,
    #   size = 3,
    #   show.legend = FALSE
    # ) +

    scale_linetype_manual(
      values = c(
        "All Expressed" = "solid",
        "All DEGs" = "solid",
        "Shared DEGs" = "longdash",
        "Species-Specific DEGs" = "dotdash",
        "Non-DEGs" = "dotted"
      )
    ) +

    scale_color_manual(
      values = colors
    ) +

    scale_linewidth_manual(
      values = c(
        "All Expressed" = 1.8,
        "All DEGs" = 1.8,
        "Shared DEGs" = 1.2,
        "Species-Specific DEGs" = 1.2,
        "Non-DEGs" = 1.2
      )
    ) +

    labs(
      title = "Inter-species correlation peaks at mid-development",
      x = "Differentiation Transition",
      y = expression(R^2),
      linetype = "Dataset",
      linewidth = "Dataset",
      color = "Dataset"
    ) +

    coord_cartesian(
      ylim = c(0, 1)
    ) +

    theme_minimal(
      base_size = 14
    ) +
    
    guides(
  color = "none",
  linetype = guide_legend(title = "Line Type"),
  linewidth = "none"
) +

    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(
        angle = 0,
        hjust = 0.5
      ),
      plot.title = element_text(face = "bold")
    )
}
```

## RLCPM_G1_4SpTp

### Gene Expression and DGE

#### All Genes

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_AllExpressed, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/AllExpressedGenes_logcpm_G1_SpTp_All_NoRep.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("All Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_comparisons <- c(
  "Day02",
  "Day24",
  "Day45",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons
)

saveRDS(matched_DEGs,"data/DGE/MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

matched_deg_genes <- readRDS("data/DGE/MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")
all_expressed_genes <- readRDS("data/DGE/AllExpressedGenes_logcpm_G1_SpTp_All_NoRep.RDS")


Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("Non-DEG at any transition - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

```

#### Plotting R2 Together

```{r Gene_FC_Quadrants_RLCOMP_G1_4SpTp_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkblue",
  "Shared DEGs" = "darkblue",
  "Species-Specific DEGs" = "darkblue",
  "Non-DEGs" = "darkblue"
)

plot_R2_trajectories(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```





### Transcription Factors

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% TF_list) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TF_AllExpressed_logcpm_G1_SpTp_All_NoRep.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% TF_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("All TF Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

trajectory_comparisons <- c(
  "Day02",
  "Day24",
  "Day45",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  TF_list
)

saveRDS(matched_DEGs,"data/DGE/TF_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

matched_deg_genes <- readRDS("data/DGE/TF_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("TF Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

all_expressed_genes <- readRDS("data/DGE/TF_AllExpressed_logcpm_G1_SpTp_All_NoRep.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("TF Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/TF_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

```

#### Plotting R2 Together

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/TF_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkgreen",
  "Shared DEGs" = "darkgreen",
  "Species-Specific DEGs" = "darkgreen",
  "Non-DEGs" = "darkgreen"
)

plot_R2_trajectories(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```




### KZFPs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_list) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_AllExpressed_logcpm_G1_SpTp_All_NoRep.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")
comparisons <- c("H_Day02", "H_Day24", "H_Day45", "H_Day515", "H_Day1530","C_Day02", "C_Day24", "C_Day45", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("All KZFP Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

trajectory_comparisons <- c(
  "Day02",
  "Day24",
  "Day45",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  KZFPs_list
)

saveRDS(matched_DEGs,"data/DGE/KZFP_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

matched_deg_genes <- readRDS("data/DGE/KZFP_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("KZFP Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "KZFP Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory Non-DEGs

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep.RDS")

all_expressed_genes <- readRDS("data/DGE/KZFP_AllExpressed_logcpm_G1_SpTp_All_NoRep.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("KZFP Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day24", "Day45", "Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day24
plots$Day45
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/KZFP_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

```

#### Plotting R2 Together

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/KZFP_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkred",
  "Shared DEGs" = "darkred",
  "Species-Specific DEGs" = "darkred",
  "Non-DEGs" = "darkred"
)

plot_R2_trajectories(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```


## RLCPM_G1_4SpTp No4

### Gene Expression and DGE

#### All Genes

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/AllExpressedGenes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons
)

saveRDS(matched_DEGs,"data/DGE/MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")
all_expressed_genes <- readRDS("data/DGE/AllExpressedGenes_logcpm_G1_SpTp_All_NoRep_No4.RDS")


Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("Non-DEG at any transition - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")
```

#### Plotting R2 Together

```{r Genes_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkblue",
  "Shared DEGs" = "darkblue",
  "Species-Specific DEGs" = "darkblue",
  "Non-DEGs" = "darkblue"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```



### Transcription Factors

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% TF_list) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TF_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% TF_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All TF Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

TF_df <- readRDS("data/TEs/TF_df.RDS")

TF_list <- TF_df$Gene

trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  TF_list
)

saveRDS(matched_DEGs,"data/DGE/TF_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/TF_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("TF Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/TF_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

all_expressed_genes <- readRDS("data/DGE/TF_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("TF Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25","Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/TF_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

```

#### Plotting R2 Together

```{r TF_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/TF_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkgreen",
  "Shared DEGs" = "darkgreen",
  "Species-Specific DEGs" = "darkgreen",
  "Non-DEGs" = "darkgreen"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```




### KZFPs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_list) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All KZFP Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r KZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_list <- KZFPs$Label


trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  KZFPs_list
)

saveRDS(matched_DEGs,"data/DGE/KZFP_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/KZFP_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("KZFP Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

all_expressed_genes <- readRDS("data/DGE/KZFP_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("KZFP Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25","Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/KZFP_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

```

#### Plotting R2 Together

```{r ZFP_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/KZFP_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkred",
  "Shared DEGs" = "darkred",
  "Species-Specific DEGs" = "darkred",
  "Non-DEGs" = "darkred"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```
### KZFP G100 mya

```{r KZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r KZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_G100 <- KZFPs %>% 
  filter(`KZFP age` > 100)

KZFPs_G100_list <- unique(KZFPs_G100$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_G100_list) %>% 
    pull(Symbol)
)
#61

saveRDS(gene_set,"data/DGE/KZFP_G100_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r KZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_G100 <- KZFPs %>% 
  filter(`KZFP age` > 100)

KZFPs_G100_list <- unique(KZFPs_G100$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_G100_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_G100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All KZFP Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r KZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_G100 <- KZFPs %>% 
  filter(`KZFP age` > 100)

KZFPs_G100_list <- unique(KZFPs_G100$Label)


trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  KZFPs_G100_list
)

saveRDS(matched_DEGs,"data/DGE/KZFP_G100_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r ZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_G100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/KZFP_G100_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("KZFP Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r ZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_G100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

all_expressed_genes <- readRDS("data/DGE/KZFP_G100_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("KZFP Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25","Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/KZFP_G100_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

```

#### Plotting R2 Together

```{r ZFP_G100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/KZFP_G100_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkred",
  "Shared DEGs" = "darkred",
  "Species-Specific DEGs" = "darkred",
  "Non-DEGs" = "darkred"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```




### KZFP 50_100 mya

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_50_100 <- KZFPs %>% 
  filter(`KZFP age` >= 50, `KZFP age` < 100)

KZFPs_50_100_list <- unique(KZFPs_50_100$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_50_100_list) %>% 
    pull(Symbol)
)
#61

saveRDS(gene_set,"data/DGE/KZFP_50_100_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_50_100 <- KZFPs %>% 
  filter(`KZFP age` >= 50, `KZFP age` < 100)

KZFPs_50_100_list <- unique(KZFPs_50_100$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_50_100_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_50_100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All KZFP Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_50_100 <- KZFPs %>% 
  filter(`KZFP age` >= 50, `KZFP age` < 100)

KZFPs_50_100_list <- unique(KZFPs_50_100$Label)


trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  KZFPs_50_100_list
)

saveRDS(matched_DEGs,"data/DGE/KZFP_50_100_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_50_100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/KZFP_50_100_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("KZFP Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_50_100_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

all_expressed_genes <- readRDS("data/DGE/KZFP_50_100_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("KZFP Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25","Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/KZFP_50_100_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

```

#### Plotting R2 Together

```{r ZFP_50_100_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/KZFP_50_100_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkred",
  "Shared DEGs" = "darkred",
  "Species-Specific DEGs" = "darkred",
  "Non-DEGs" = "darkred"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```


### KZFP L50 mya

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_EmptyTable, fig.width=7, fig.height=5}

R2_table <- data.frame(
  Gene_Set = character(),
  Transition = character(),
  R2 = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)
```

#### All Expressed

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_AllExpressed, fig.width=7, fig.height=5}

TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_L50 <- KZFPs %>% 
  filter(`KZFP age` < 50)

KZFPs_L50_list <- unique(KZFPs_L50$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_L50_list) %>% 
    pull(Symbol)
)
#61

saveRDS(gene_set,"data/DGE/KZFP_L50_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All Expressed Genes - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Expressed Genes"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530


R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Expressed Genes",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  ))


```


#### All Trajectory DEGs

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_DEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
comparisons <- c("H_Day02", "H_Day25", "H_Day515", "H_Day1530","C_Day02", "C_Day25", "C_Day515", "C_Day1530")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_L50 <- KZFPs %>% 
  filter(`KZFP age` < 50)

KZFPs_L50_list <- unique(KZFPs_L50$Label)

gene_set <- unique(
  TopTable %>%
    filter(Comparisons %in% comparisons) %>%
    filter(Symbol %in% KZFPs_L50_list) %>% 
    filter(adj.P.Val < 0.05) %>% 
    pull(Symbol)
)

saveRDS(gene_set,"data/DGE/KZFP_L50_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("All KZFP Trajectory DEGs - with Symbol IDs")
length(gene_set)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = gene_set,
  genes_label = "All Trajectory DEGs"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "All Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Matched Trajectory DEGs

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_MatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

KZFPs <- readRDS("data/TEs/KZFPs.RDS")

KZFPs_L50 <- KZFPs %>% 
  filter(`KZFP age` < 50)

KZFPs_L50_list <- unique(KZFPs_L50$Label)


trajectory_comparisons <- c(
  "Day02",
  "Day25",
  "Day515",
  "Day1530"
)

matched_DEGs <- get_matched_degs(
  TopTable = TopTable,
  transitions = trajectory_comparisons,
  genes_to_include =  KZFPs_L50_list
)

saveRDS(matched_DEGs,"data/DGE/KZFP_L50_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

print("Trajectory DEGs matched across species in at least one transition - with Symbol IDs")
length(matched_DEGs)


plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = matched_DEGs,
  genes_label = "Species-Shared Trajectory DEGs \n (at least on Transition)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-Shared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```

#### All Trajectory DEGs (NonShared)

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonMatchedDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_L50_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

matched_deg_genes <- readRDS("data/DGE/KZFP_L50_MatchedTrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

umatched_deg_genes <- setdiff(trajectory_deg_genes, matched_deg_genes)

print("KZFP Trajectory DEGs with no species matches at any transition - with Symbol IDs")
length(umatched_deg_genes)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25", "Day515", "Day1530"),
  genes_to_include = umatched_deg_genes,
  genes_label = "TF Species-NonShared Trajectory DEGs \n (no Transitions)"
)


plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Species-NonShared Trajectory DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

```
#### All Trajectory Non-DEGs

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_NonDEGs, fig.width=7, fig.height=5}
TopTable <- readRDS("data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

trajectory_deg_genes <- readRDS("data/DGE/KZFP_L50_TrajectoryDEGs_logcpm_G1_SpTp_All_NoRep_No4.RDS")

all_expressed_genes <- readRDS("data/DGE/KZFP_L50_AllExpressed_logcpm_G1_SpTp_All_NoRep_No4.RDS")

Non_DEGs <- setdiff(
  all_expressed_genes,
  trajectory_deg_genes
)

print("KZFP Non-DEG at all transitions - with Symbol IDs")
length(Non_DEGs)

plots <- plot_species_comparison(
  TopTable = TopTable,
  comps = c("Day02", "Day25","Day515", "Day1530"),
  genes_to_include = Non_DEGs,
  genes_label = "Trajectory Non-DEGs \n (no Transitions)"
)

plots$Day02
plots$Day25
plots$Day515
plots$Day1530

R2_table <- rbind(
  R2_table,
  data.frame(
    Gene_Set = "Trajectory Non-DEGs",
    Transition = names(plots),
    R2 = sapply(plots, `[[`, "r2"),
    P_value = sapply(plots, `[[`, "pval")
  )
)

saveRDS(R2_table,"data/DGE/KZFP_L50_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

```

#### Plotting R2 Together

```{r ZFP_L50_FC_Quadrants_RLCOMP_G1_4SpTp_No4_Plotting, fig.width=7, fig.height=6}

TrajectoryR2Genes <- readRDS("data/DGE/KZFP_L50_TrajectoryR2Genes_logcpm_G1_SpTp_All_NoRep_No4.RDS")

RNA_colors <- c(
  "All Expressed" = "black",
  "All DEGs" = "darkred",
  "Shared DEGs" = "darkred",
  "Species-Specific DEGs" = "darkred",
  "Non-DEGs" = "darkred"
)

plot_R2_trajectories_No4(
  R2_table = TrajectoryR2Genes,
  colors = RNA_colors
)

```

```{r Updating Website, eval = FALSE, fig.height=10, fig.width=10}

# git -> commit all changes
# git -> push
wflow_publish("analysis/RNA_CorrelationPlots.Rmd")

```


