---
title: "TETranscript_Chimp"
output:
  html_document:
    toc: true
    toc_float: true
date: "2026-08-18"
---


```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

## TE Expression

### Setting enviroment

#### Loading libraries

```{r Enviroment Set Up, }
#loading libraries
library("tidyverse")
library("rmarkdown")
library("readxl")
library("readr")
library("ggrepel")
library("ggfortify")
library("patchwork")
library("eulerr")
library("org.Hs.eg.db")
library("AnnotationDbi")
library("ggVennDiagram")
library("ggrastr")
library("edgeR")
library("limma")
library("ggsignif")
library("dplyr")
library("purrr")
library("ggplot2")
library("tidyr")
library("AnnotationDbi")
library("org.Hs.eg.db")
library("ggalluvial")
library("tibble")
library("pheatmap")
library("Cormotif")
library("workflowr")
library("ComplexHeatmap")

# BiocManager::install("ggalluvial")

```

#### Defining functions

```{r FischerTest_OddsRation_plotting}
plot_OR <- function(results, group_to_plot, ylim = c(-2, 2)) {

  plot_data <- results %>%
    filter(Comparison_Group %in% group_to_plot) %>%
    mutate(
      label = ifelse(FDR < 0.05, "*", ""),
      log2_OR = log2(Odds_Ratio)
    )

  ggplot(plot_data, aes(Comparison, Odds_Ratio)) +
    geom_point(size = 3) +
    geom_text(
      aes(label = label),
      vjust = -1
    ) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    coord_cartesian(ylim = ylim) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

#FischerTest_fate_OddsRation_plotting

plot_Fate_OR <- function(results, ylim = c(0, 2)) {

  plot_data <- results %>%
    mutate(
      label = ifelse(FDR < 0.05, "*", ""),
      log2_OR = log2(Odds_Ratio)
    )

  ggplot(plot_data, aes(x = Comparison, y = Odds_Ratio)) +
    geom_point(size = 3) +
    geom_text(
      aes(label = label),
      vjust = -1
    ) +
    geom_hline(
      yintercept = 1,
      linetype = "dashed"
    ) +
    facet_wrap(~Fate) +
    coord_cartesian(ylim = ylim) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      x = "HvC developmental comparison",
      y = "Odds Ratio",
      title = "KZFP enrichment by HvC DEG fate"
    )
}

#Volcano_Plot_Function

generate_volcano_plot <- function(toptable, title) {
  
  # #check for entrezid 
  # if(!"Entrez_ID" %in% colnames(toptable)) stop("Entrez_ID col not present")
  # 
  #make significance labels
  
  toptable <- toptable %>% 
    mutate(Significance = case_when(
      logFC > 0 & adj.P.Val < 0.05 ~ "Upregulated",
      logFC < 0 & adj.P.Val < 0.05 ~ "Downregulated",
      TRUE  ~ "Not Significant"
    ))
  
  #factor significance
  toptable$Significance <- factor(
    toptable$Significance, 
    levels = c("Upregulated", 
               "Not Significant", 
               "Downregulated")
  )
  
  #count genes in each category
  upgenes <- sum(toptable$Significance == "Upregulated")
  nsgenes <- sum(toptable$Significance == "Not Significant")
  downgenes <- sum(toptable$Significance == "Downregulated")
  
  
  #labels for legend
  legend_lab <- c(
    paste0("Upregulated: ", upgenes),
    paste0("Not Significant: ", nsgenes),
    paste0("Downregulated: ", downgenes)
  )
  
  #colors 
  color_map <- c("Upregulated" = "blue",
                 "Not Significant" = "grey30",
                 "Downregulated" = "red")
  
  #generate volcano plots
  p <- ggplot(toptable, aes(x = logFC, 
                            y = -log10(P.Value),
                            color = Significance)) +
    geom_point_rast(alpha = 0.8, size = 3) +
    scale_color_manual(values = color_map, 
                       labels = legend_lab,
                       breaks = c("Upregulated",
                                  "Not Significant", 
                                  "Downregulated"))  +
    xlim(-20,20) +
    labs(title = title, 
         x = expression("log"[2]*"FC"),
         y = expression("-log"[10]*"p-value")) +
    # theme_custom() +
    theme(legend.position = "right")
  
  return(p)
}

#FC_Plot_Function_sig_trajectory

plot_cluster_fc_sig_trajectory <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(logFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$logFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = logFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("logFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG FC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(-2, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}

#FC_Plot_Function_sig_trajectory_No4

plot_cluster_fc_sig_trajectory_No4 <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(logFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$logFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = logFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("logFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG FC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(-2, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}

#absFC_Plot_Function_sig_trajectory_No4

plot_cluster_absfc_sig_trajectory <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(absFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$absFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = absFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("absFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG absFC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(0, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}

#absFC_Plot_Function_sig_trajectory_No4

plot_cluster_absfc_sig_trajectory_No4 <- function(gene_vector, cluster_name, top_table) {

  df <- top_table %>%
    dplyr::filter(Gene %in% gene_vector) %>%
    mutate(Comparisons = factor(Comparisons, levels = c(
      "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
    )))

  cat(cluster_name, ":", nrow(df), "rows,",
      length(unique(df$Gene)), "unique genes\n")

  # ✅ Skip if no data
  if (nrow(df) == 0) {
    message(cluster_name, ": no data — skipping")
    return(NULL)
  }

  # ✅ Skip if only one comparison present
  if (length(unique(df$Comparisons)) < 2) {
    message(cluster_name, ": only one comparison — skipping stats")
    return(NULL)
  }

  # Stats
  kw <- kruskal.test(absFC ~ Comparisons, data = df)

  pw_kw <- pairwise.wilcox.test(
    df$absFC,
    df$Comparisons,
    p.adjust.method = "BH"
  )

  print(pw_kw)

  p <- ggplot(df, aes(x = Comparisons, y = absFC)) +
    geom_boxplot(aes(fill = Comparisons)) +
    theme_bw() +
    xlab(" ") +
    ylab("absFC") +
    ggtitle(
      paste0(
        cluster_name,
        " Trajectroy DEG absFC Comparisons\nKruskal–Wallis p = ",
        signif(kw$p.value, 3)
      )
    ) +
    coord_cartesian(ylim = c(0, 2)) +
    theme(
      plot.title = element_text(size = rel(1.3), hjust = 0.5),
      axis.title = element_text(size = 15),
      axis.text.x = element_text(size = 8, angle = 15)
    )

  return(p)
}


#get_directional_genes_HvC1

get_fc_directions <- function(traj_label, coef_name, comparison_label,
                                 top_table_all, fit_obj, p_cutoff = 0.05) {

  # ---------------------------
  # DE genes
  # ---------------------------
  tt <- topTable(fit_obj, coef = coef_name, number = Inf)

  sig_tt <- tt %>%
    dplyr::filter(adj.P.Val < p_cutoff) %>%
    tibble::rownames_to_column("Gene") %>%
    dplyr::distinct(Gene, .keep_all = TRUE)

  # ---------------------------
  # Trajcetory FC data
  # ---------------------------
  df_fc <- top_table_all %>%
    dplyr::filter(Comparisons == comparison_label) %>%
    dplyr::distinct(Gene, .keep_all = TRUE)

  # ---------------------------
  # classify direction
  # ---------------------------
  df_combo <- df_fc %>%
  dplyr::inner_join(sig_tt, by = "Gene") %>%
  dplyr::mutate(
    direction = dplyr::case_when(
      logFC.x > 0 ~ "Increased",
      logFC.x < 0 ~ "Decreased",
      TRUE ~ "No_Change"
    ),
    magnitude = abs(logFC.x)
  )

  # ---------------------------
  # output
  # ---------------------------
  list(
    all = df_combo,
    Increased = dplyr::filter(df_combo, direction == "Increased"),
    Decreased = dplyr::filter(df_combo, direction == "Decreased"),
    n_sig = nrow(sig_tt)
  )
}
```

```{r FischerTest_OddsRation_plotting}

plot_TE_proportions <- function(
    file,
    level = "TE_Class",
    direction = NULL
) {
  
  direction_results <- readRDS(file)
  
  # Separate TE information
  direction_results <- lapply(direction_results, function(x) {
    
    x[c("all", "Increased", "Decreased")] <- lapply(
      x[c("all", "Increased", "Decreased")],
      function(df) {
        tidyr::separate(
          df,
          Gene,
          into = c("TE_Name", "TE_Family", "TE_Class"),
          sep = ":",
          remove = FALSE
        )
      }
    )
    
    x
  })
  
  # Combine "all" results
  direction_all <- lapply(direction_results, `[[`, "all")
  
  direction_all <- dplyr::bind_rows(
  direction_all,
  .id = "Trajectory"
) %>%
  dplyr::filter(!grepl("\\?", TE_Class))
  
  # Filter direction
  if (!is.null(direction)) {
    direction_all <- direction_all %>%
      dplyr::filter(direction == !!direction)
  }
  
  # Trajectory order
  trajectory_levels <- c(
    "Traj02_genes",
    "Traj25_genes",
    "Traj515_genes",
    "Traj1530_genes"
  )
  
  direction_all <- direction_all %>%
    dplyr::mutate(
      Trajectory = factor(
        Trajectory,
        levels = trajectory_levels
      )
    )
  
  # ----------------------------
  # TE Class
  # ----------------------------
  
  if (level == "TE_Class") {
    
    proportions <- direction_all %>%
      dplyr::count(Trajectory, TE_Class) %>%
      dplyr::group_by(Trajectory) %>%
      dplyr::mutate(Proportion = n / sum(n)) %>%
      dplyr::ungroup()
    
    p <- ggplot(
      proportions,
      aes(x = Trajectory, y = Proportion, fill = TE_Class)
    ) +
      geom_col() +
      scale_y_continuous(labels = scales::percent) +
      labs(
        x = NULL,
        y = "Proportion of TE DEGs",
        fill = "TE Class"
      ) +
      theme_classic()
    
    return(p)
  }
  
  
  # ----------------------------
  # TE Family
  # One plot for each Class
  # ----------------------------
  
  if (level == "TE_Family") {
    
    class_list <- split(
      direction_all,
      direction_all$TE_Class
    )
    
    plots <- lapply(class_list, function(df) {
      
      proportions <- df %>%
        dplyr::count(Trajectory, TE_Family) %>%
        dplyr::group_by(Trajectory) %>%
        dplyr::mutate(Proportion = n / sum(n)) %>%
        dplyr::ungroup()
      
      ggplot(
        proportions,
        aes(x = Trajectory, y = Proportion, fill = TE_Family)
      ) +
        geom_col() +
        scale_y_continuous(labels = scales::percent) +
        labs(
          title = unique(df$TE_Class),
          x = NULL,
          y = "Proportion of TE DEGs",
          fill = "TE Family"
        ) +
        theme_classic()
    })
    
    return(plots)
  }
  
  
  # ----------------------------
  # TE Name
  # One plot for each Class-Family
  # ----------------------------
  
  if (level == "TE_Name") {
    
    group_list <- split(
      direction_all,
      list(
        direction_all$TE_Class,
        direction_all$TE_Family
      ),
      drop = TRUE
    )
    
    plots <- lapply(group_list, function(df) {
      
      proportions <- df %>%
        dplyr::count(Trajectory, TE_Name) %>%
        dplyr::group_by(Trajectory) %>%
        dplyr::mutate(Proportion = n / sum(n)) %>%
        dplyr::ungroup()
      
      ggplot(
        proportions,
        aes(x = Trajectory, y = Proportion, fill = TE_Name)
      ) +
        geom_col() +
        scale_y_continuous(labels = scales::percent) +
        labs(
          title = paste(
            unique(df$TE_Class),
            unique(df$TE_Family),
            sep = " - "
          ),
          x = NULL,
          y = "Proportion of TE DEGs",
          fill = "TE Name"
        ) +
        theme_classic()
    })
    
    return(plots)
  }
  
  
  stop("level must be 'TE_Class', 'TE_Family', or 'TE_Name'")
}
```

#### Data Upload

```{r TopTable_Symbols}
TopTable <- readRDS("data/DGE/TopTable_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")

# Extract Ensembl IDs for every row
ens_ids <- sub("\\..*$", "", TopTable$Gene)

# Make a lookup table
symbol_map <- mapIds(
    org.Hs.eg.db,
    keys = unique(ens_ids),
    column = "SYMBOL",
    keytype = "ENSEMBL",
    multiVals = "first"
)

TopTable$Symbol <- unname(symbol_map[ens_ids])

# saveRDS(TopTable,"data/DGE/TopTable_Symbols_Filt_logcpm_G1_SpTp_RNA_fc_NoRep_No4.RDS")
```

```{r Chimp_TE_Expression_Data_Upload}

sample_names <- readLines("data/TEs/TE_Expression/chimp_samples.txt")
# 
# count_tables <- lapply(sample_names, function(sample) {
#   counts <- read.table(
#     paste0("data/TEs/TE_Expression/", sample, ".cntTable"),
#     header = TRUE,
#     row.names = 1
#   )
# 
#   colnames(counts) <- sample
# 
#   counts
# })
# 
# combined_counts <- Reduce(
#   function(x, y) {
#     x <- data.frame(Feature = rownames(x), x, check.names = FALSE)
#     y <- data.frame(Feature = rownames(y), y, check.names = FALSE)
# 
#     merged <- merge(x, y, by = "Feature", all = TRUE)
# 
#     rownames(merged) <- merged$Feature
#     merged$Feature <- NULL
# 
#     merged
#   },
#   count_tables
# )
# 
# saveRDS(combined_counts,"data/TEs/TE_Expression/Chimp/chimp_TECount_fc.RDS")

```

### TECount

#### Filtering Steps

##### Log2cpm G1 >= 4 individuals in >= 1SpTp

```{r Chimp_TE_Expression_Filtering}

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata_full <- metadata_combined  %>% 
    filter(Species == "C")

counts <- readRDS("data/TEs/TE_Expression/Chimp/chimp_TECount_fc.RDS")

colnames(counts)[match(metadata_full$Library_ID, colnames(counts))] <-metadata_full$SampleName

fc_full <- counts

print("Number of Repeats + Genes")
length(unique(rownames(fc_full)))

##---------------------##
##---------tes---------##
##---------------------##

metadata <- metadata_full

fc <- fc_full %>% 
  filter(grepl(":", rownames(.)))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_repeats_fc.RDS")

print("Number of Repeats")
length(unique(rownames(fc )))

fc <- fc_full %>% 
  filter(grepl(":", rownames(.))) %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", rownames(.)))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")

print("Number of TEs (LINE|SINE|LTR|DNA|Retroposon)")
length(unique(rownames(fc)))

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>% 
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})


fc_filtered <- fc[keep, ]

# saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")

print("Number of TEs after expression filtering")
length(unique(rownames(fc_filtered)))

##---------------------##
##-------tes_no4-------##
##---------------------##

fc <- fc_full %>% 
  filter(grepl(":", rownames(.))) %>% 
  dplyr::select(-contains("_Day4"))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_repeats_fc_no4.RDS")

fc <- fc_full %>% 
  filter(grepl(":", rownames(.))) %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", rownames(.))) %>% 
  dplyr::select(-contains("_Day4"))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>% 
    filter(Timepoint != "Day4") %>% 
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})


fc_filtered <- fc[keep, ]

# saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")

print("Number of TEs after expression filtering excluding day 4")
length(unique(rownames(fc_filtered)))

##---------------------##
##--------genes--------##
##---------------------##

metadata <- metadata_full

fc <- fc_full %>% 
  filter(!grepl(":", rownames(.)))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_genes_fc.RDS")

print("Number of Genes")
length(unique(rownames(fc)))

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>%
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})


fc_filtered <- fc[keep, ]

# saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_genes_fc_filtered.RDS")

print("Number of Genes after expression filtering")
length(unique(rownames(fc_filtered)))

##---------------------##
##------genes_No4------##
##---------------------##

fc <- fc_full %>% 
  filter(!grepl(":", rownames(.))) %>% 
  dplyr::select(-contains("_Day4"))

# saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4.RDS")

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>% 
    filter(Timepoint != "Day4") %>% 
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})


fc_filtered <- fc[keep, ]

# saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4_filtered.RDS")

print("Number of Genes after expression filtering excluding day 4")
length(unique(rownames(fc_filtered)))






```

#### TEs

##### Stats

###### All days

```{r Chimp_TE_Expression_Filtering_3}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )
```

###### All filtered

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames\nFiltered cpm > 2\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nFiltered cpm > 2\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nFiltered cpm > 2\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nFiltered cpm > 2\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No day 4

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepName\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nNo Day4",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nNo Day4",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No 4 filtered

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames\nNo Day4 Filtered cpm > 2\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nNo Day4 Filtered cpm > 2 \n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nNo Day4 Filtered cpm > 2\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nNo Day4 Filtered cpm > 2\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

##### Cor_HeatMap

###### All days

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation"
)

```

###### All filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")
metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nFiltered cpm > 2\n4 Ind at least 1 SpTp"
)

```

###### No day 4

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nNo Day4"
)

```

###### No 4 filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nNo Day4 Filtered cpm > 2\n4 Ind at least 1 SpTp"
)

```


##### PCA

###### All days

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


###### All filtered

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nFiltered cpm > 2\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nFiltered cpm > 2\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No day 4

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  arrange(match(SampleName, colnames(log2cpm))) %>% 
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No 4 filtered
 
```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  arrange(match(SampleName, colnames(log2cpm))) %>% 
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nNo day 4 Filtered cpm > 2\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nNo day 4 Filtered cpm > 2\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

##### DGE

###### All days

```{r Chimp_TE_DGE_Differential_Data}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

# counts_norep <- fc %>%
#   dplyr::select(-(contains("R")))
# 
# metadata <- metadata_combined %>%
#   filter(Species == "C") %>%
#   filter(Individual != "H4R")
# 
# 
# metadata$dgelist <- c(rep(c("Chimp_Day0","Chimp_Day2","Chimp_Day4","Chimp_Day5","Chimp_Day15","Chimp_Day30"),7))
# 
# #create DGEList object
# dge <- DGEList(counts = counts_norep)
# dge$samples$group <- factor(metadata$dgelist)
# dge <- calcNormFactors(dge, method = "TMM")
# dge$samples
# 
# saveRDS(dge,"data/TEs/TE_Expression/Chimp/dge_chimp_tes.RDS")
# 
# dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_chimp_tes.RDS")
# 
# dim(dge)
# dge$samples
# 
# metadata$Timepoint <- factor(
#   metadata$Timepoint,
#   levels = c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")
# )
# 
# design_int <- model.matrix(
#   ~ Timepoint,
#   data = metadata
# )
# 
# colnames(design_int)
# colnames(design_int) <- make.names(colnames(design_int))
# 
# saveRDS(design_int,"data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")
# 
# # Step 1: voom first
# v1 <- voom(dge, design_int, plot = TRUE)
# 
# saveRDS(v1,"data/TEs/TE_Expression/Chimp/v1_chimp_tes.RDS")
# 
# # Step 2: estimate duplicate correlation on voom-transformed data
# corfit <- duplicateCorrelation(v1, design_int, block = metadata$Individual)
# corfit.correlation <- corfit$consensus.correlation
# 
# # Step 3: voom again including correlation
# v2 <- voom(dge, design_int, plot = TRUE,
#            block = metadata$Individual,
#            correlation = corfit.correlation)
# 
# saveRDS(v2,"data/TEs/TE_Expression/Chimp/v2_chimp_tes.RDS")
# 
# # Step 4: fit linear model
# fit <- lmFit(v2, design_int, block = metadata$Individual, correlation = corfit.correlation)
# fit <- eBayes(fit)
# 
# saveRDS(fit,"data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")


design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")

dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_chimp_tes.RDS")

v1 <- readRDS("data/TEs/TE_Expression/Chimp/v1_chimp_tes.RDS")

v2 <- readRDS("data/TEs/TE_Expression/Chimp/v2_chimp_tes.RDS")

fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")

pal <- rainbow(length(unique(metadata$Timepoint)))
day <- metadata$Timepoint
plotDensities(v2, col = pal[as.numeric(day)], legend = FALSE)

chimp_idx <- metadata$Species == "C"

pal <- rainbow(length(unique(metadata$Timepoint)))


# Chimp
plotDensities(
  v2[, chimp_idx],
  col = pal[as.numeric(as.factor(metadata$Timepoint[chimp_idx]))],
  main = "",
  legend = FALSE,
  axes = FALSE
)

axis(1)
axis(2)
mtext("log2 CPM", side = 1, line = 2)

title("Chimp")
  legend(
      "right",
      legend = metadata$Individual[chimp_idx],  # or Timepoint, etc.
      col = rainbow(sum(chimp_idx)),
      lty = 1,
     cex = 0.7
    )

species_levels <- unique(metadata$Species)
day_levels <- unique(metadata$Timepoint)

for (sp in species_levels) {
  for (day in day_levels) {

    idx <- metadata$Species == sp &
           metadata$Timepoint == day

    if (sum(idx) > 0) {

      cols <- rainbow(sum(idx))

      plotDensities(
        v2[, idx],
        col = cols,
        main = paste(sp, day),
        legend = FALSE,
        axes = FALSE
      )

      axis(1)
      axis(2)
      mtext("log2 CPM", side = 1, line = 2)

      legend(
        "right",
        legend = metadata$Individual[idx],
        col = cols,
        lty = 1,
        cex = 0.7
      )
    }
  }
}


```

```{r Chimp_TE_Expression_Mode}
fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")
design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")


contrast_matrix <- makeContrasts(
  Day2_vs_Day0 = TimepointDay2,
  Day4_vs_Day2 = TimepointDay4 - TimepointDay2,
  Day5_vs_Day4 = TimepointDay5 - TimepointDay4,
  Day15_vs_Day5 = TimepointDay15 - TimepointDay5,
  Day30_vs_Day15 = TimepointDay30 - TimepointDay15,
  levels = design_int
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

plotSA(fit2, main = "RNA Model")

results_summary <- decideTests(fit2,
                               adjust.method = "BH",
                               p.value = 0.05)
summary(results_summary)

# saveRDS(fit2,"data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")
```

```{r Chimp_TE_Expression_TopTable}
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")

TopTables_names <- colnames(fit2$coefficients) 

TopTable <- do.call(
  rbind,
  lapply(TopTables_names,function(ct) {
    
    tt <- topTable(
      fit2,
      coef = ct,
      number = Inf,
      sort.by = "p"
    )
    
    #add identifiers
    tt$Gene <- rownames(tt)
    tt$Comparisons <- ct
    
    tt
  })
)

ComparisonNames <- unique(TopTable$Comparisons)

# saveRDS(TopTable,"data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
# saveRDS(ComparisonNames,"data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

head(TopTable)
```

```{r Chimp_TE_Expression_VolcancoPlots}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

# Calculate global axis limits
x_limits <- range(TopTable$logFC, na.rm = TRUE)
y_limits <- range(-log10(TopTable$P.Value), na.rm = TRUE)

plot <- list()

for (n in ComparisonNames) {
  data_n <- TopTable %>% 
    dplyr::filter(Comparisons == n)
  
  plot[[n]] <- generate_volcano_plot(data_n, n) +
    ggplot2::xlim(x_limits) +
    ggplot2::ylim(y_limits)
}

# Print plots
for (n in ComparisonNames) {
  print(plot[[n]])
}
```

```{r Chimp_TE_Expression_Subsetting0.05}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

RNA_sets <- map(
  ComparisonNames,
  ~ {
    df <- TopTable %>%
      filter(Comparisons == paste0(.x))

    list(
      DEGs = df %>% filter(adj.P.Val < 0.05),
      NonDEGs = df %>% filter(adj.P.Val >= 0.05)
    )
  }
)


names(RNA_sets) <- paste0(ComparisonNames)

# saveRDS(RNA_sets, "data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")
```

```{r Chimp_TE_Expression_FC, fig.height=6, fig.width=10}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")

TopTable$absFC <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

Traj02_genes <- RNA_sets$Day2_vs_Day0$DEGs$Gene
Traj24_genes <- RNA_sets$Day4_vs_Day2$DEGs$Gene
Traj45_genes <- RNA_sets$Day5_vs_Day4$DEGs$Gene
Traj515_genes <- RNA_sets$Day15_vs_Day5$DEGs$Gene
Traj1530_genes <- RNA_sets$Day30_vs_Day15$DEGs$Gene
all_exp_TEs <- unique(TopTable$Gene)

length(Traj02_genes)
length(Traj24_genes)
length(Traj45_genes)
length(Traj515_genes)
length(Traj1530_genes)
length(all_exp_TEs)

cluster_lists <- list(
 Traj02_genes  = Traj02_genes,
  Traj24_genes  = Traj24_genes,
  Traj45_genes  = Traj45_genes,
  Traj515_genes  = Traj515_genes,
  Traj1530_genes  = Traj1530_genes,
  all_exp_TEs = all_exp_TEs
)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_absfc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

```


```{r Chimp_TE_Expression_Directionality}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")

traj_info <- list(
  Traj02_genes = list(coef = "Day2_vs_Day0", comp = "Day2_vs_Day0"),
  Traj24_genes = list(coef = "Day4_vs_Day2", comp = "Day4_vs_Day2"),
  Traj24_genes = list(coef = "Day5_vs_Day4", comp = "Day5_vs_Day4"),
  Traj515_genes = list(coef ="Day15_vs_Day5", comp = "Day15_vs_Day5"),
Traj1530_genes = list(coef = "Day30_vs_Day15", comp = "Day30_vs_Day15")
)

direction_results <- setNames(
  lapply(names(traj_info), function(name) {

    info <- traj_info[[name]]

    get_fc_directions(
      traj_label = name,
      coef_name = info$coef,
      comparison_label = info$comp,
      top_table_all = TopTable,
      fit_obj = fit2
    )

  }),
  names(traj_info)
)

# saveRDS(direction_results, "data/TEs/TE_Expression/Chimp/Directions_chimp_tes.RDS")

```

```{r Chimp_TE_Expression_Directionality}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
direction_results <- readRDS("data/TEs/TE_Expression/Chimp/Directions_chimp_tes.RDS")

TopTable$absfc <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

aa <- unique(direction_results$Traj02_genes$all$Gene)
ab <- unique(direction_results$Traj02_genes$Increased$Gene)
ac <- unique(direction_results$Traj02_genes$Decreased$Gene)


ba <- unique(direction_results$Traj24_genes$all$Gene)
bb <- unique(direction_results$Traj24_genes$Increased$Gene)
bc <- unique(direction_results$Traj24_genes$Decreased$Gene)

ca <- unique(direction_results$Traj45_genes$all$Gene)
cb <- unique(direction_results$Traj45_genes$Increased$Gene)
cc <- unique(direction_results$Traj45_genes$Decreased$Gene)

da <- unique(direction_results$Traj515_genes$all$Gene)
db <- unique(direction_results$Traj515_genes$Increased$Gene)
dc <- unique(direction_results$Traj515_genes$Decreased$Gene)



ea <- unique(direction_results$Traj1530_genes$all$Gene)
eb <- unique(direction_results$Traj1530_genes$Increased$Gene)
ec <- unique(direction_results$Traj1530_genes$Decreased$Gene)


cluster_lists <- list(
  Traj02_All = aa,
  Traj02_Increased = ab,
  Traj02_Decrease = ac,

  
  Traj24_All = ba,
  Traj24_Increased = bb,
  Traj24_Decrease = bc,

  Traj45_All = ca,
  Traj45_Increased = cb,
  Traj45_Decrease = cc,
  
  Traj515_All = da,
  Traj515_ChimpPos = db,
  Traj515_Decrease = dc,

 
  Traj1530_All = ea,
  Traj1530_Increased = eb,
  Traj1530_Decrease = ec

)


cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})

names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)


```

```{r Chimp_TE_Expression_Directionality_results}

direction_results <- readRDS("data/TEs/TE_Expression/Chimp/Directions_chimp_tes.RDS")

direction_results <- lapply(direction_results, function(x) {
  
  x[c("all", "Increased", "Decreased")] <- lapply(
    x[c("all", "Increased", "Decreased")],
    function(df) {
      tidyr::separate(
        df,
        Gene,
        into = c("TE_Name", "TE_Family", "TE_Class"),
        sep = ":",
        remove = FALSE
      )
    }
  )
  
  x
})


direction_all <- lapply(direction_results, `[[`, "all")

direction_all <- dplyr::bind_rows(
  direction_all,
  .id = "Trajectory"
)

class_proportions_all <- direction_all %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions_all,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()

class_proportions <- direction_all %>%
  filter(direction == "Increased") %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of Increased TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()


class_proportions <- direction_all %>%
  filter(direction == "Decreased") %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of Decreased TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()
```

###### No Day4
```{r Chimp_TE_DGE_Differential_Data}
fc <- readRDS("data/TEs/TE_Expression/Chimp/Chimp_tes_fc_no4_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

counts_norep <- fc %>%
  dplyr::select(-(contains("R"))) %>% 
  dplyr::select(-contains("_Day4"))

metadata <- metadata_combined %>%
  filter(Species == "H") %>%
  filter(Individual != "H4R") %>% 
  filter(Timepoint != "Day4")


metadata$dgelist <- c(rep(c("Chimp_Day0","Chimp_Day2","Chimp_Day5","Chimp_Day15","Chimp_Day30"),7))

# #create DGEList object
# dge <- DGEList(counts = counts_norep)
# dge$samples$group <- factor(metadata$dgelist)
# dge <- calcNormFactors(dge, method = "TMM")
# dge$samples
# 
# saveRDS(dge,"data/TEs/TE_Expression/Chimp/dge_Chimp_tes_no4.RDS")
# 
# dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_Chimp_tes_no4.RDS")
# 
# dim(dge)
# dge$samples
# 
# metadata$Timepoint <- factor(
#   metadata$Timepoint,
#   levels = c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")
# )
# 
# design_int <- model.matrix(
#   ~ Timepoint,
#   data = metadata
# )
# 
# colnames(design_int)
# colnames(design_int) <- make.names(colnames(design_int))
# 
# saveRDS(design_int,"data/TEs/TE_Expression/Chimp/design_Chimp_tes_no4.RDS")
# 
# # Step 1: voom first
# v1 <- voom(dge, design_int, plot = TRUE)
# 
# saveRDS(v1,"data/TEs/TE_Expression/Chimp/v1_Chimp_tes_no4.RDS")
# 
# # Step 2: estimate duplicate correlation on voom-transformed data
# corfit <- duplicateCorrelation(v1, design_int, block = metadata$Individual)
# corfit.correlation <- corfit$consensus.correlation
# 
# # Step 3: voom again including correlation
# v2 <- voom(dge, design_int, plot = TRUE,
#            block = metadata$Individual,
#            correlation = corfit.correlation)
# 
# saveRDS(v2,"data/TEs/TE_Expression/Chimp/v2_Chimp_tes_no4.RDS")
# 
# # Step 4: fit linear model
# fit <- lmFit(v2, design_int, block = metadata$Individual, correlation = corfit.correlation)
# fit <- eBayes(fit)
# 
# saveRDS(fit,"data/TEs/TE_Expression/Chimp/fit_Chimp_tes_no4.RDS")


design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_Chimp_tes_no4.RDS")

dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_Chimp_tes_no4.RDS")

v1 <- readRDS("data/TEs/TE_Expression/Chimp/v1_Chimp_tes_no4.RDS")

v2 <- readRDS("data/TEs/TE_Expression/Chimp/v2_Chimp_tes_no4.RDS")

fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_Chimp_tes_no4.RDS")

pal <- rainbow(length(unique(metadata$Timepoint)))
day <- metadata$Timepoint
plotDensities(v2, col = pal[as.numeric(day)], legend = FALSE)

Chimp_idx <- metadata$Species == "H"

pal <- rainbow(length(unique(metadata$Timepoint)))


# Chimp
plotDensities(
  v2[, Chimp_idx],
  col = pal[as.numeric(as.factor(metadata$Timepoint[Chimp_idx]))],
  main = "",
  legend = FALSE,
  axes = FALSE
)

axis(1)
axis(2)
mtext("log2 CPM", side = 1, line = 2)

title("Chimp")
  legend(
      "right",
      legend = metadata$Individual[Chimp_idx],  # or Timepoint, etc.
      col = rainbow(sum(Chimp_idx)),
      lty = 1,
     cex = 0.7
    )

species_levels <- unique(metadata$Species)
day_levels <- unique(metadata$Timepoint)

for (sp in species_levels) {
  for (day in day_levels) {

    idx <- metadata$Species == sp &
           metadata$Timepoint == day

    if (sum(idx) > 0) {

      cols <- rainbow(sum(idx))

      plotDensities(
        v2[, idx],
        col = cols,
        main = paste(sp, day),
        legend = FALSE,
        axes = FALSE
      )

      axis(1)
      axis(2)
      mtext("log2 CPM", side = 1, line = 2)

      legend(
        "right",
        legend = metadata$Individual[idx],
        col = cols,
        lty = 1,
        cex = 0.7
      )
    }
  }
}


```

```{r Chimp_TE_Expression_Mode}
fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_Chimp_tes_no4.RDS")
design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_Chimp_tes_no4.RDS")


contrast_matrix <- makeContrasts(
  Day2_vs_Day0 = TimepointDay2,
  Day5_vs_Day2 = TimepointDay5 - TimepointDay2,
  Day15_vs_Day5 = TimepointDay15 - TimepointDay5,
  Day30_vs_Day15 = TimepointDay30 - TimepointDay15,
  levels = design_int
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

plotSA(fit2, main = "RNA Model")

results_summary <- decideTests(fit2,
                               adjust.method = "BH",
                               p.value = 0.05)
summary(results_summary)

# saveRDS(fit2,"data/TEs/TE_Expression/Chimp/fit2_Chimp_tes_no4.RDS")
```

```{r Chimp_TE_Expression_TopTable}
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_Chimp_tes_no4.RDS")

TopTables_names <- colnames(fit2$coefficients) 

TopTable <- do.call(
  rbind,
  lapply(TopTables_names,function(ct) {
    
    tt <- topTable(
      fit2,
      coef = ct,
      number = Inf,
      sort.by = "p"
    )
    
    #add identifiers
    tt$Gene <- rownames(tt)
    tt$Comparisons <- ct
    
    tt
  })
)

ComparisonNames <- unique(TopTable$Comparisons)

# saveRDS(TopTable,"data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
# saveRDS(ComparisonNames,"data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")

head(TopTable)
```

```{r Chimp_TE_Expression_VolcancoPlots}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")

# Calculate global axis limits
x_limits <- range(TopTable$logFC, na.rm = TRUE)
y_limits <- range(-log10(TopTable$P.Value), na.rm = TRUE)

plot <- list()

for (n in ComparisonNames) {
  data_n <- TopTable %>% 
    dplyr::filter(Comparisons == n)
  
  plot[[n]] <- generate_volcano_plot(data_n, n) +
    ggplot2::xlim(x_limits) +
    ggplot2::ylim(y_limits)
}

# Print plots
for (n in ComparisonNames) {
  print(plot[[n]])
}
```

```{r Chimp_TE_Expression_Subsetting0.05}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")

RNA_sets <- map(
  ComparisonNames,
  ~ {
    df <- TopTable %>%
      filter(Comparisons == paste0(.x))

    list(
      DEGs = df %>% filter(adj.P.Val < 0.05),
      NonDEGs = df %>% filter(adj.P.Val >= 0.05)
    )
  }
)


names(RNA_sets) <- paste0(ComparisonNames)

# saveRDS(RNA_sets, "data/TEs/TE_Expression/Chimp/RNASets_Chimp_tes_no4.RDS")
```

```{r Chimp_TE_Expression_FC, fig.height=6, fig.width=10}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_Chimp_tes_no4.RDS")

TopTable$absFC <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
  ))

Traj02_genes <- RNA_sets$Day2_vs_Day0$DEGs$Gene
Traj25_genes <- RNA_sets$Day5_vs_Day2$DEGs$Gene
Traj515_genes <- RNA_sets$Day15_vs_Day5$DEGs$Gene
Traj1530_genes <- RNA_sets$Day30_vs_Day15$DEGs$Gene
all_exp_TEs <- unique(TopTable$Gene)

length(Traj02_genes)
length(Traj25_genes)
length(Traj515_genes)
length(Traj1530_genes)
length(all_exp_TEs)

cluster_lists <- list(
 Traj02_genes  = Traj02_genes,
  Traj25_genes  = Traj25_genes,
  Traj515_genes  = Traj515_genes,
  Traj1530_genes  = Traj1530_genes,
  all_exp_TEs = all_exp_TEs
)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory_No4(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_absfc_sig_trajectory_No4(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

```


```{r Chimp_TE_Expression_Directionality}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_Chimp_tes_no4.RDS")
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_Chimp_tes_no4.RDS")

traj_info <- list(
  Traj02_genes = list(coef = "Day2_vs_Day0", comp = "Day2_vs_Day0"),
  Traj25_genes = list(coef = "Day5_vs_Day2", comp = "Day5_vs_Day2"),
  Traj515_genes = list(coef ="Day15_vs_Day5", comp = "Day15_vs_Day5"),
Traj1530_genes = list(coef = "Day30_vs_Day15", comp = "Day30_vs_Day15")
)

direction_results <- setNames(
  lapply(names(traj_info), function(name) {

    info <- traj_info[[name]]

    get_fc_directions(
      traj_label = name,
      coef_name = info$coef,
      comparison_label = info$comp,
      top_table_all = TopTable,
      fit_obj = fit2
    )

  }),
  names(traj_info)
)

# saveRDS(direction_results, "data/TEs/TE_Expression/Chimp/Directions_Chimp_tes_no4.RDS")

```

```{r Chimp_TE_Expression_Directionality}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_Chimp_tes_no4.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_Chimp_tes_no4.RDS")
direction_results <- readRDS("data/TEs/TE_Expression/Chimp/Directions_Chimp_tes_no4.RDS")

TopTable$absfc <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day5_vs_Day2","Day15_vs_Day5","Day30_vs_Day15"
  ))

aa <- unique(direction_results$Traj02_genes$all$Gene)
ab <- unique(direction_results$Traj02_genes$Increased$Gene)
ac <- unique(direction_results$Traj02_genes$Decreased$Gene)


ba <- unique(direction_results$Traj25_genes$all$Gene)
bb <- unique(direction_results$Traj25_genes$Increased$Gene)
bc <- unique(direction_results$Traj25_genes$Decreased$Gene)

da <- unique(direction_results$Traj515_genes$all$Gene)
db <- unique(direction_results$Traj515_genes$Increased$Gene)
dc <- unique(direction_results$Traj515_genes$Decreased$Gene)



ea <- unique(direction_results$Traj1530_genes$all$Gene)
eb <- unique(direction_results$Traj1530_genes$Increased$Gene)
ec <- unique(direction_results$Traj1530_genes$Decreased$Gene)


cluster_lists <- list(
  Traj02_All = aa,
  Traj02_Increased = ab,
  Traj02_Decrease = ac,

  
  Traj25_All = ba,
  Traj25_Increased = bb,
  Traj25_Decrease = bc,

  
  Traj515_All = da,
  Traj515_ChimpPos = db,
  Traj515_Decrease = dc,

 
  Traj1530_All = ea,
  Traj1530_Increased = eb,
  Traj1530_Decrease = ec

)


cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory_No4(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})

names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)


```




###### Proportion Plots

```{r Chimp_TE_Expression_Directionality_results}

class_plot <- plot_TE_proportions( "data/TEs/TE_Expression/Chimp/Directions_Chimp_tes_no4.RDS", level = "TE_Class" ) 

class_plot

```


```{r Chimp_TE_Expression_Directionality_results}

family_plots <- plot_TE_proportions(
  "data/TEs/TE_Expression/Chimp/Directions_Chimp_tes_no4.RDS",
  level = "TE_Family"
)

for (p in family_plots) {
  print(p)
}
```


```{r Chimp_TE_Expression_Directionality_results, fig.width=10, fig.height=10}
name_plots <- plot_TE_proportions(
  "data/TEs/TE_Expression/Chimp/Directions_Chimp_tes_no4.RDS",
  level = "TE_Name"
)

for (p in name_plots) {
  print(p)
}
```

#### Genes

##### Stats

###### All days

```{r Chimp_TE_Expression_Filtering_3, fig.width=10, fig.height=5}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of Genes",xaxt = "n", xlab= "", ylim = c(0,1200))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of Genes",xaxt = "n", xlab= "", ylim = c(0,60))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of Genes",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of Genes",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )
```

###### All filtered

```{r Chimp_TE_Expression_Filtering_3, fig.width=10, fig.height=5}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of Genes\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1200))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of Genes\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,60))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of Genes\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of Genes\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No day 4

```{r Chimp_TE_Expression_Filtering_3, fig.width=10, fig.height=5}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of Genes\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,1200))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of Genes\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,60))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of Genes\nNo Day4",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of Genes\nNo Day4",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No 4 filtered

```{r Chimp_TE_Expression_Filtering_3, fig.width=10, fig.height=5}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of Genes\nNo Day4\ - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1200))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of Genes\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,60))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of Genes\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of Genes\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

##### Cor_HeatMap

###### All days

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.8, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Gene Expression Sample-Sample Correlation"
)

```

###### All filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_filtered.RDS")
metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.8, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Gene Expression Sample-Sample Correlation\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
)

```

###### No day 4

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.8, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Gene Expression Sample-Sample Correlation\nNo Day4"
)

```

###### No 4 filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.8, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Gene Expression Sample-Sample Correlation\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
)

```
##### PCA

###### All days

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nScaled"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


###### All filtered

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nScaled\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No day 4

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nScaled\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No 4 filtered
 
```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nNo day 4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp Gene Expression PC1–PC2\nScaled\nNo day 4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```
##### DGE

```{r Chimp_TE_DGE_Differential_Data}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_genes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

# counts_norep <- fc %>%
#   dplyr::select(-(contains("R")))
# 
# metadata <- metadata_combined %>%
#   filter(Species == "C") %>%
#   filter(Individual != "H4R")
# 
# 
# metadata$dgelist <- c(rep(c("Chimp_Day0","Chimp_Day2","Chimp_Day4","Chimp_Day5","Chimp_Day15","Chimp_Day30"),7))
# 
# #create DGEList object
# dge <- DGEList(counts = counts_norep)
# dge$samples$group <- factor(metadata$dgelist)
# dge <- calcNormFactors(dge, method = "TMM")
# dge$samples
# # 
# saveRDS(dge,"data/TEs/TE_Expression/Chimp/dge_chimp_genes.RDS")
# 
# dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_chimp_genes.RDS")
# 
# dim(dge)
# dge$samples
# 
# metadata$Timepoint <- factor(
#   metadata$Timepoint,
#   levels = c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")
# )
# 
# design_int <- model.matrix(
#   ~ Timepoint,
#   data = metadata
# )
# 
# colnames(design_int)
# colnames(design_int) <- make.names(colnames(design_int))
# 
# saveRDS(design_int,"data/TEs/TE_Expression/Chimp/design_chimp_genes.RDS")
# 
# # # Step 1: voom first
# v1 <- voom(dge, design_int, plot = TRUE)
# 
# saveRDS(v1,"data/TEs/TE_Expression/Chimp/v1_chimp_genes.RDS")
# 
# # Step 2: estimate duplicate correlation on voom-transformed data
# corfit <- duplicateCorrelation(v1, design_int, block = metadata$Individual)
# corfit.correlation <- corfit$consensus.correlation
# 
# # Step 3: voom again including correlation
# v2 <- voom(dge, design_int, plot = TRUE,
#            block = metadata$Individual,
#            correlation = corfit.correlation)
# 
# saveRDS(v2,"data/TEs/TE_Expression/Chimp/v2_chimp_genes.RDS")
# 
# 
# 
# 
# # Step 4: fit linear model
# fit <- lmFit(v2, design_int, block = metadata$Individual, correlation = corfit.correlation)
# fit <- eBayes(fit)
# # 
# saveRDS(fit,"data/TEs/TE_Expression/Chimp/fit_chimp_genes.RDS")

design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_genes.RDS")

fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_genes.RDS")

v1 <- readRDS("data/TEs/TE_Expression/Chimp/v1_chimp_genes.RDS")

v2 <- readRDS("data/TEs/TE_Expression/Chimp/v2_chimp_genes.RDS")

pal <- rainbow(length(unique(metadata$Timepoint)))
day <- metadata$Timepoint
plotDensities(v2, col = pal[as.numeric(day)], legend = FALSE)

chimp_idx <- metadata$Species == "C"

pal <- rainbow(length(unique(metadata$Timepoint)))


# Chimp
plotDensities(
  v2[, chimp_idx],
  col = pal[as.numeric(as.factor(metadata$Timepoint[chimp_idx]))],
  main = "",
  legend = FALSE,
  axes = FALSE
)

axis(1)
axis(2)
mtext("log2 CPM", side = 1, line = 2)

title("Chimp")
  legend(
      "right",
      legend = metadata$Individual[chimp_idx],  # or Timepoint, etc.
      col = rainbow(sum(chimp_idx)),
      lty = 1,
     cex = 0.7
    )

species_levels <- unique(metadata$Species)
day_levels <- unique(metadata$Timepoint)

for (sp in species_levels) {
  for (day in day_levels) {

    idx <- metadata$Species == sp &
           metadata$Timepoint == day

    if (sum(idx) > 0) {

      cols <- rainbow(sum(idx))

      plotDensities(
        v2[, idx],
        col = cols,
        main = paste(sp, day),
        legend = FALSE,
        axes = FALSE
      )

      axis(1)
      axis(2)
      mtext("log2 CPM", side = 1, line = 2)

      legend(
        "right",
        legend = metadata$Individual[idx],
        col = cols,
        lty = 1,
        cex = 0.7
      )
    }
  }
}


```

```{r Chimp_TE_Expression_Mode}
fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_genes.RDS")
design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_genes.RDS")


contrast_matrix <- makeContrasts(
  Day2_vs_Day0 = TimepointDay2,
  Day4_vs_Day2 = TimepointDay4 - TimepointDay2,
  Day5_vs_Day4 = TimepointDay5 - TimepointDay4,
  Day15_vs_Day5 = TimepointDay15 - TimepointDay5,
  Day30_vs_Day15 = TimepointDay30 - TimepointDay15,
  levels = design_int
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

plotSA(fit2, main = "RNA Model")

results_summary <- decideTests(fit2,
                               adjust.method = "BH",
                               p.value = 0.05)
summary(results_summary)

# saveRDS(fit2,"data/TEs/TE_Expression/Chimp/fit2_chimp_genes.RDS")
```

```{r Chimp_TE_Expression_TopTable}
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_genes.RDS")

TopTables_names <- colnames(fit2$coefficients) 

TopTable <- do.call(
  rbind,
  lapply(TopTables_names,function(ct) {
    
    tt <- topTable(
      fit2,
      coef = ct,
      number = Inf,
      sort.by = "p"
    )
    
    #add identifiers
    tt$Gene <- rownames(tt)
    tt$Comparisons <- ct
    
    tt
  })
)

ComparisonNames <- unique(TopTable$Comparisons)

# saveRDS(TopTable,"data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
# saveRDS(ComparisonNames,"data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")

head(TopTable)
```

```{r Chimp_TE_Expression_VolcancoPlots}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")

# Calculate global axis limits
x_limits <- range(TopTable$logFC, na.rm = TRUE)
y_limits <- range(-log10(TopTable$P.Value), na.rm = TRUE)

plot <- list()

for (n in ComparisonNames) {
  data_n <- TopTable %>% 
    dplyr::filter(Comparisons == n)
  
  plot[[n]] <- generate_volcano_plot(data_n, n) +
    ggplot2::xlim(x_limits) +
    ggplot2::ylim(y_limits)
}

# Print plots
for (n in ComparisonNames) {
  print(plot[[n]])
}
```

```{r Chimp_TE_Expression_Subsetting0.05}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")

RNA_sets <- map(
  ComparisonNames,
  ~ {
    df <- TopTable %>%
      filter(Comparisons == paste0(.x))

    list(
      DEGs = df %>% filter(adj.P.Val < 0.05),
      NonDEGs = df %>% filter(adj.P.Val >= 0.05)
    )
  }
)


names(RNA_sets) <- paste0(ComparisonNames)

# saveRDS(RNA_sets, "data/TEs/TE_Expression/Chimp/RNASets_chimp_genes.RDS")
```

```{r Chimp_TE_Expression_FC, fig.height=6, fig.width=10}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_genes.RDS")

TopTable$absFC <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

Traj02_genes <- RNA_sets$Day2_vs_Day0$DEGs$Gene
Traj24_genes <- RNA_sets$Day4_vs_Day2$DEGs$Gene
Traj45_genes <- RNA_sets$Day5_vs_Day4$DEGs$Gene
Traj515_genes <- RNA_sets$Day15_vs_Day5$DEGs$Gene
Traj1530_genes <- RNA_sets$Day30_vs_Day15$DEGs$Gene
all_exp_TEs <- unique(TopTable$Gene)

length(Traj02_genes)
length(Traj24_genes)
length(Traj45_genes)
length(Traj515_genes)
length(Traj1530_genes)
length(all_exp_TEs)

cluster_lists <- list(
 Traj02_genes  = Traj02_genes,
  Traj24_genes  = Traj24_genes,
  Traj45_genes  = Traj45_genes,
  Traj515_genes  = Traj515_genes,
  Traj1530_genes  = Traj1530_genes,
  all_exp_TEs = all_exp_TEs
)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_absfc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

```


```{r Chimp_TE_Expression_Directionality}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_genes.RDS")
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_genes.RDS")

traj_info <- list(
  Traj02_genes = list(coef = "Day2_vs_Day0", comp = "Day2_vs_Day0"),
  Traj24_genes = list(coef = "Day4_vs_Day2", comp = "Day4_vs_Day2"),
  Traj24_genes = list(coef = "Day5_vs_Day4", comp = "Day5_vs_Day4"),
  Traj515_genes = list(coef ="Day15_vs_Day5", comp = "Day15_vs_Day5"),
Traj1530_genes = list(coef = "Day30_vs_Day15", comp = "Day30_vs_Day15")
)

direction_results <- setNames(
  lapply(names(traj_info), function(name) {

    info <- traj_info[[name]]

    get_fc_directions(
      traj_label = name,
      coef_name = info$coef,
      comparison_label = info$comp,
      top_table_all = TopTable,
      fit_obj = fit2
    )

  }),
  names(traj_info)
)

# saveRDS(direction_results, "data/TEs/TE_Expression/Chimp/Directions_chimp_genes.RDS")

```

```{r Chimp_TE_Expression_Directionality}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_genes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_genes.RDS")
direction_results <- readRDS("data/TEs/TE_Expression/Chimp/Directions_chimp_genes.RDS")

TopTable$absfc <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

aa <- unique(direction_results$Traj02_genes$all$Gene)
ab <- unique(direction_results$Traj02_genes$Increased$Gene)
ac <- unique(direction_results$Traj02_genes$Decreased$Gene)


ba <- unique(direction_results$Traj24_genes$all$Gene)
bb <- unique(direction_results$Traj24_genes$Increased$Gene)
bc <- unique(direction_results$Traj24_genes$Decreased$Gene)

ca <- unique(direction_results$Traj45_genes$all$Gene)
cb <- unique(direction_results$Traj45_genes$Increased$Gene)
cc <- unique(direction_results$Traj45_genes$Decreased$Gene)

da <- unique(direction_results$Traj515_genes$all$Gene)
db <- unique(direction_results$Traj515_genes$Increased$Gene)
dc <- unique(direction_results$Traj515_genes$Decreased$Gene)



ea <- unique(direction_results$Traj1530_genes$all$Gene)
eb <- unique(direction_results$Traj1530_genes$Increased$Gene)
ec <- unique(direction_results$Traj1530_genes$Decreased$Gene)


cluster_lists <- list(
  Traj02_All = aa,
  Traj02_Increased = ab,
  Traj02_Decrease = ac,

  
  Traj24_All = ba,
  Traj24_Increased = bb,
  Traj24_Decrease = bc,

  Traj45_All = ca,
  Traj45_Increased = cb,
  Traj45_Decrease = cc,
  
  Traj515_All = da,
  Traj515_ChimpPos = db,
  Traj515_Decrease = dc,

 
  Traj1530_All = ea,
  Traj1530_Increased = eb,
  Traj1530_Decrease = ec

)


cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})

names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)


```

### Instances_Table

#### Count Tables

```{r Chimp_TE_Instances_Expression_Data_Upload}
sample_names <- readLines("data/TEs/TE_Expression/chimp_samples.txt")

RNA_Metadata <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

# metadata <- RNA_Metadata %>% 
#   filter(Species == "C")
# 
# # Read all TEinstance files
# count_tables <- lapply(sample_names, function(sample) {
#   
#   counts <- read.table(
#     paste0(
#       "data/TEs/TECount_Instances/",
#       sample,
#       ".TEinstance_counts.txt"
#     ),
#     header = TRUE,
#     row.names = 1,
#     check.names = FALSE
#   )
#   
#   # Keep annotations + count
#   counts <- counts[, 1:5, drop = FALSE]
#   
#   # Rename the count column to the sample name
#   colnames(counts)[5] <- sample
#   
#   counts
# })
# 
# # Start with annotations + first sample
# combined_counts_Chimp_annotated <- count_tables[[1]]
# 
# # Add only the count column from each remaining sample
# for (i in 2:length(count_tables)) {
#   
#   combined_counts_Chimp_annotated <- merge(
#     combined_counts_Chimp_annotated,
#     count_tables[[i]][, 5, drop = FALSE],
#     by = "row.names",
#     all = TRUE
#   )
#   
#   rownames(combined_counts_Chimp_annotated) <- combined_counts_Chimp_annotated$Row.names
#   combined_counts_Chimp_annotated$Row.names <- NULL
# }
# 
# saveRDS(combined_counts_Chimp_annotated,"data/TEs/TECount_Instances/Chimp/chimp_TEInstances_fc.RDS")
# 
# 
# dim(combined_counts_Chimp_annotated)
# head(combined_counts_Chimp_annotated)
# colnames(combined_counts_Chimp_annotated)
# 
# setequal(
#   colnames(combined_counts_Chimp_annotated)[5:ncol(combined_counts_Chimp_annotated)],
#   sample_names
# )


```


#### Filtering Steps

##### Log2cpm G1 >= 4 individuals in >= 1SpTp

```{r Chimp_TE_Expression_Filtering}

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata_full <- metadata_combined  %>% 
    filter(Species == "C")

counts <- readRDS("data/TEs/TECount_Instances/Chimp/chimp_TEInstances_fc.RDS")

fc_full <- counts

print("Number of Repeats")
length(unique(rownames(fc_full)))

##---------------------##
##---------tes---------##
##---------------------##

fc <- fc_full %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", family_id)) %>% 
  dplyr::select(-matches("gene_id|family_id|class_id|strand"))

fc_annotated <- fc_full %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", family_id))

colnames(fc)[match(metadata_full$Library_ID, colnames(fc))] <-metadata_full$SampleName

saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_TEInstances_fc.RDS")

print("Number of TEs (LINE|SINE|LTR|DNA|Retroposon)")
length(unique(rownames(fc)))

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>% 
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})

fc_filtered <- fc_annotated[keep, ]

saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_TEInstances_fc_filtered_annotated.RDS")

print("Number of TEs after expression filtering")
length(unique(rownames(fc_filtered)))

##---------------------##
##-------tes_no4-------##
##---------------------##


metadata <- metadata_full %>% 
    filter(Timepoint != "Day4")

fc <- fc_full %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", family_id)) %>% 
  dplyr::select(-matches("gene_id|family_id|class_id|strand")) %>% 
  dplyr::select(-contains("_Day4"))

fc_annotated <- fc_full %>% 
  filter(grepl("LINE|SINE|LTR|DNA|Retroposon", family_id)) %>% 
  dplyr::select(-contains("_Day4"))

colnames(fc)[match(metadata_full$Library_ID, colnames(fc))] <-metadata_full$SampleName

saveRDS(fc,"data/TEs/TE_Expression/Chimp/chimp_TEInstances_fc_No4.RDS")

print("Number of TEs (LINE|SINE|LTR|DNA|Retroposon) No Day 4")
length(unique(rownames(fc)))

fc_filtering <- fc %>%
  dplyr::select(-contains("R"))

cpm_values <- edgeR::cpm(
  fc_filtering,
  log = FALSE
)

metadata <- metadata_full %>% 
    dplyr::select(-contains("R"))
  
group <- metadata$Cond

keep <- apply(cpm_values > 2, 1, function(x) {
  any(tapply(x, group, sum) >= 4)
})

fc_filtered <- fc_annotated[keep, ]

saveRDS(fc_filtered,"data/TEs/TE_Expression/Chimp/chimp_TEInstances_fc_no4_filtered_annotated.RDS")

print("Number of TEs after expression filtering excluding day 4")
length(unique(rownames(fc_filtered)))

```

#### TEs

##### Stats

###### All days

```{r Chimp_TE_Expression_Filtering_3}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )
```

###### All filtered

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No day 4

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepName\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nNo Day4",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nNo Day4",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nNo Day4",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

###### No 4 filtered

```{r Chimp_TE_Expression_Filtering_3}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")


metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

cpm <- cpm(fc,log=FALSE)

log2cpm <- cpm(fc,log=TRUE)

col_labels <- metadata$SampleName

boxplot(fc, main = "Boxplots of Chimp count of TE RepNames\nNo Day4\ - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,1000))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

boxplot(cpm, main = "Boxplots of Chimp cpm of TE RepNames\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "", ylim = c(0,500))
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size


boxplot(log2cpm, main = "Boxplots of Chimp Log2cpm of TE RepNames\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",xaxt = "n", xlab= "")
axis(1,
     at = 1:length(col_labels),  # positions (one per sample)
     labels = col_labels,        # your labels vector
     las = 2,               # rotate text vertically (like srt=90)
     cex.axis = 0.6)        # shrink label size

hist(log2cpm, main = "Histogram of Chimp Log2cpm of TE RepNames\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp",
     xlab =expression("Log"[2]*" counts-per-million"), col =5 )

```

##### Cor_HeatMap

###### All days

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation"
)

```

###### All filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")
metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
)

```

###### No day 4

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nNo Day4"
)

```

###### No 4 filtered

```{r Chimp_TE_Expression_Cor, fig.height=10,fig.width=10}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C") %>% 
    filter(Timepoint != "Day4")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

colnames(fc) <- metadata$SampleName[
  match(colnames(fc), metadata$Library_ID)
]

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

######Cor_HeatMap####
Cor_log2cpm <- cor(log2cpm, method = "spearman")
individual <- metadata$Individual
species <- metadata$Species
timepoint <- metadata$Timepoint
timepoint <- factor(timepoint,levels = c("Day0","Day2","Day4","Day5","Day15","Day30"))
Cor_metadata <- data.frame(
  sample_cor = colnames(Cor_log2cpm),
  species_cor = species,
  timepoint_cor = timepoint,
  individual_cor = individual
)


rownames(Cor_metadata) <- Cor_metadata$sample_cor

library(RColorBrewer)

my_breaks <- seq(0.9, 1, length.out = 101)
my_colors <- colorRampPalette(c("darkblue","blue","lightblue", "white","salmon","red","darkred"))(100)

pheatmap(
  Cor_log2cpm,
  fontsize_row = 5,
  fontsize_col = 5,
  annotation_col = Cor_metadata[, c("species_cor", "timepoint_cor", "individual_cor")],
  annotation_colors = ann_colors,
  clustering_distance_rows = "correlation",
  clustering_distance_cols = "correlation",
  color = my_colors,
  breaks = my_breaks,
  main = "Chimp Repname TEs Sample-Sample Correlation\nNo Day4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
)

```


##### PCA

###### All days

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


###### All filtered

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nFiltered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No day 4

```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nNo day 4"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

###### No 4 filtered
 
```{r Chimp_TE_Expression_PCA_PC1_PC2, fig.height=6,fig.width=8}

fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_no4_filtered.RDS")
ann_colors <- readRDS("data/Raw_Data/concat/ann_Colors.RDS")
time_order <- c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")


metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

metadata <- metadata_combined  %>% 
    filter(Species == "C")

log2cpm <- cpm(fc,log=TRUE)

ann_colors$timepoint_cor <- ann_colors$timepoint_cor[time_order]

metadata <- metadata %>%
  dplyr::filter(SampleName %in% colnames(fc)) %>%
  dplyr::slice(match(colnames(fc), SampleName)) %>%
  dplyr::mutate(
    Timepoint = factor(Timepoint, levels = time_order)
  )

##NoScale

prcomp <- prcomp(t(log2cpm), scale=FALSE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nNo day 4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )



##Scale

prcomp <- prcomp(t(log2cpm), scale=TRUE, center=TRUE)
prcomp_meta <- prcomp$x %>% cbind(., metadata)
#PCA checks
#k=1

ggplot2::autoplot(
    prcomp,
    data = prcomp_meta,
    colour = "Timepoint",
    shape  = "Species",
    size   = 4,
    x = 1,
    y = 2
) +
  scale_y_reverse() +
  scale_color_manual(values = ann_colors$timepoint_cor) +
  ggrepel::geom_text_repel(
    label = prcomp_meta$CellLine,
    vjust = -0.5,
    max.overlaps = 50
  ) +
  ggtitle(
  "Chimp TE Expression PC1–PC2\nScaled\nNo day 4 - Filtered Log2cpm > 1\n4 Ind at least 1 SpTp"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

##### DGE

```{r Chimp_TE_DGE_Differential_Data}
fc <- readRDS("data/TEs/TE_Expression/Chimp/chimp_tes_fc_filtered.RDS")

metadata_combined <- readRDS("data/Raw_Data/concat/RNA_Metadata.RDS")

# counts_norep <- fc %>%
#   dplyr::select(-(contains("R")))
# 
# metadata <- metadata_combined %>% 
#   filter(Species == "C") %>%
#   filter(Individual != "H4R")
#   
# 
# metadata$dgelist <- c(rep(c("Chimp_Day0","Chimp_Day2","Chimp_Day4","Chimp_Day5","Chimp_Day15","Chimp_Day30"),7))
# 
# #create DGEList object
# dge <- DGEList(counts = counts_norep)
# dge$samples$group <- factor(metadata$dgelist)
# dge <- calcNormFactors(dge, method = "TMM")
# dge$samples
# 
saveRDS(dge,"data/TEs/TE_Expression/Chimp/dge_chimp_tes.RDS")

dge <- readRDS("data/TEs/TE_Expression/Chimp/dge_chimp_tes.RDS")

# dim(dge)
# dge$samples
# 
# metadata$Timepoint <- factor(
#   metadata$Timepoint,
#   levels = c("Day0", "Day2", "Day4", "Day5", "Day15", "Day30")
# )
# 
# design_int <- model.matrix(
#   ~ Timepoint,
#   data = metadata
# )
# 
# colnames(design_int)
# colnames(design_int) <- make.names(colnames(design_int))
# 
saveRDS(design_int,"data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")

design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")

# # Step 1: voom first
# v1 <- voom(dge, design_int, plot = TRUE)
# 
saveRDS(v1,"data/TEs/TE_Expression/Chimp/v1_chimp_tes.RDS")
# 
# # Step 2: estimate duplicate correlation on voom-transformed data
# corfit <- duplicateCorrelation(v1, design_int, block = metadata$Individual)
# corfit.correlation <- corfit$consensus.correlation
# 
# # Step 3: voom again including correlation
# v2 <- voom(dge, design_int, plot = TRUE,
#            block = metadata$Individual,
#            correlation = corfit.correlation)
# 
saveRDS(v2,"data/TEs/TE_Expression/Chimp/v2_chimp_tes.RDS")
# 
# # Step 4: fit linear model
# fit <- lmFit(v2, design_int, block = metadata$Individual, correlation = corfit.correlation)
# fit <- eBayes(fit)
# 
saveRDS(fit,"data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")

v1 <- readRDS("data/TEs/TE_Expression/Chimp/v1_chimp_tes.RDS")

v2 <- readRDS("data/TEs/TE_Expression/Chimp/v2_chimp_tes.RDS")

fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")

pal <- rainbow(length(unique(metadata$Timepoint)))
day <- metadata$Timepoint
plotDensities(v2, col = pal[as.numeric(day)], legend = FALSE)

chimp_idx <- metadata$Species == "C"

pal <- rainbow(length(unique(metadata$Timepoint)))


# Chimp
plotDensities(
  v2[, chimp_idx],
  col = pal[as.numeric(as.factor(metadata$Timepoint[chimp_idx]))],
  main = "",
  legend = FALSE,
  axes = FALSE
)

axis(1)
axis(2)
mtext("log2 CPM", side = 1, line = 2)

title("Chimp")
  legend(
      "right",
      legend = metadata$Individual[chimp_idx],  # or Timepoint, etc.
      col = rainbow(sum(chimp_idx)),
      lty = 1,
     cex = 0.7
    )

species_levels <- unique(metadata$Species)
day_levels <- unique(metadata$Timepoint)

for (sp in species_levels) {
  for (day in day_levels) {

    idx <- metadata$Species == sp &
           metadata$Timepoint == day

    if (sum(idx) > 0) {

      cols <- rainbow(sum(idx))

      plotDensities(
        v2[, idx],
        col = cols,
        main = paste(sp, day),
        legend = FALSE,
        axes = FALSE
      )

      axis(1)
      axis(2)
      mtext("log2 CPM", side = 1, line = 2)

      legend(
        "right",
        legend = metadata$Individual[idx],
        col = cols,
        lty = 1,
        cex = 0.7
      )
    }
  }
}


```

```{r Chimp_TE_Expression_Mode}
fit <- readRDS("data/TEs/TE_Expression/Chimp/fit_chimp_tes.RDS")
design_int <- readRDS("data/TEs/TE_Expression/Chimp/design_chimp_tes.RDS")


contrast_matrix <- makeContrasts(
  Day2_vs_Day0 = TimepointDay2,
  Day4_vs_Day2 = TimepointDay4 - TimepointDay2,
  Day5_vs_Day4 = TimepointDay5 - TimepointDay4,
  Day15_vs_Day5 = TimepointDay15 - TimepointDay5,
  Day30_vs_Day15 = TimepointDay30 - TimepointDay15,
  levels = design_int
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

plotSA(fit2, main = "RNA Model")

results_summary <- decideTests(fit2,
                               adjust.method = "BH",
                               p.value = 0.05)
summary(results_summary)

saveRDS(fit2,"data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")
```

```{r Chimp_TE_Expression_TopTable}
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")

TopTables_names <- colnames(fit2$coefficients) 

TopTable <- do.call(
  rbind,
  lapply(TopTables_names,function(ct) {
    
    tt <- topTable(
      fit2,
      coef = ct,
      number = Inf,
      sort.by = "p"
    )
    
    #add identifiers
    tt$Gene <- rownames(tt)
    tt$Comparisons <- ct
    
    tt
  })
)

ComparisonNames <- unique(TopTable$Comparisons)

saveRDS(TopTable,"data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
saveRDS(ComparisonNames,"data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

head(TopTable)
```

```{r Chimp_TE_Expression_VolcancoPlots}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

# Calculate global axis limits
x_limits <- range(TopTable$logFC, na.rm = TRUE)
y_limits <- range(-log10(TopTable$P.Value), na.rm = TRUE)

plot <- list()

for (n in ComparisonNames) {
  data_n <- TopTable %>% 
    dplyr::filter(Comparisons == n)
  
  plot[[n]] <- generate_volcano_plot(data_n, n) +
    ggplot2::xlim(x_limits) +
    ggplot2::ylim(y_limits)
}

# Print plots
for (n in ComparisonNames) {
  print(plot[[n]])
}
```

```{r Chimp_TE_Expression_Subsetting0.05}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")

RNA_sets <- map(
  ComparisonNames,
  ~ {
    df <- TopTable %>%
      filter(Comparisons == paste0(.x))

    list(
      DEGs = df %>% filter(adj.P.Val < 0.05),
      NonDEGs = df %>% filter(adj.P.Val >= 0.05)
    )
  }
)


names(RNA_sets) <- paste0(ComparisonNames)

saveRDS(RNA_sets, "data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")
```

```{r Chimp_TE_Expression_FC, fig.height=6, fig.width=10}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")

TopTable$absFC <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

Traj02_genes <- RNA_sets$Day2_vs_Day0$DEGs$Gene
Traj24_genes <- RNA_sets$Day4_vs_Day2$DEGs$Gene
Traj45_genes <- RNA_sets$Day5_vs_Day4$DEGs$Gene
Traj515_genes <- RNA_sets$Day15_vs_Day5$DEGs$Gene
Traj1530_genes <- RNA_sets$Day30_vs_Day15$DEGs$Gene
all_exp_TEs <- unique(TopTable$Gene)

length(Traj02_genes)
length(Traj24_genes)
length(Traj45_genes)
length(Traj515_genes)
length(Traj1530_genes)
length(all_exp_TEs)

cluster_lists <- list(
 Traj02_genes  = Traj02_genes,
  Traj24_genes  = Traj24_genes,
  Traj45_genes  = Traj45_genes,
  Traj515_genes  = Traj515_genes,
  Traj1530_genes  = Traj1530_genes,
  all_exp_TEs = all_exp_TEs
)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_absfc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})
names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)

```


```{r Chimp_TE_Expression_Directionality}
TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
RNA_sets <- readRDS("data/TEs/TE_Expression/Chimp/RNASets_chimp_tes.RDS")
fit2 <- readRDS("data/TEs/TE_Expression/Chimp/fit2_chimp_tes.RDS")

traj_info <- list(
  Traj02_genes = list(coef = "Day2_vs_Day0", comp = "Day2_vs_Day0"),
  Traj24_genes = list(coef = "Day4_vs_Day2", comp = "Day4_vs_Day2"),
  Traj24_genes = list(coef = "Day5_vs_Day4", comp = "Day5_vs_Day4"),
  Traj515_genes = list(coef ="Day15_vs_Day5", comp = "Day15_vs_Day5"),
Traj1530_genes = list(coef = "Day30_vs_Day15", comp = "Day30_vs_Day15")
)

direction_results <- setNames(
  lapply(names(traj_info), function(name) {

    info <- traj_info[[name]]

    get_fc_directions(
      traj_label = name,
      coef_name = info$coef,
      comparison_label = info$comp,
      top_table_all = TopTable,
      fit_obj = fit2
    )

  }),
  names(traj_info)
)

saveRDS(direction_results, "data/TEs/TE_Expression/Directions_chimp_tes.RDS")

```

```{r Chimp_TE_Expression_Directionality}

TopTable <- readRDS("data/TEs/TE_Expression/Chimp/TopTable_chimp_tes.RDS")
ComparisonNames <- readRDS("data/TEs/TE_Expression/Chimp/ComparisonNames_chimp_tes.RDS")
direction_results <- readRDS("data/TEs/TE_Expression/Directions_chimp_tes.RDS")

TopTable$absfc <- abs(TopTable$logFC)

TopTable_Traj <- TopTable %>% 
  filter(Comparisons %in% c(
    "Day2_vs_Day0","Day4_vs_Day2","Day5_vs_Day4","Day15_vs_Day5","Day30_vs_Day15"
  ))

aa <- unique(direction_results$Traj02_genes$all$Gene)
ab <- unique(direction_results$Traj02_genes$Increased$Gene)
ac <- unique(direction_results$Traj02_genes$Decreased$Gene)


ba <- unique(direction_results$Traj24_genes$all$Gene)
bb <- unique(direction_results$Traj24_genes$Increased$Gene)
bc <- unique(direction_results$Traj24_genes$Decreased$Gene)

ca <- unique(direction_results$Traj45_genes$all$Gene)
cb <- unique(direction_results$Traj45_genes$Increased$Gene)
cc <- unique(direction_results$Traj45_genes$Decreased$Gene)

da <- unique(direction_results$Traj515_genes$all$Gene)
db <- unique(direction_results$Traj515_genes$Increased$Gene)
dc <- unique(direction_results$Traj515_genes$Decreased$Gene)



ea <- unique(direction_results$Traj1530_genes$all$Gene)
eb <- unique(direction_results$Traj1530_genes$Increased$Gene)
ec <- unique(direction_results$Traj1530_genes$Decreased$Gene)


cluster_lists <- list(
  Traj02_All = aa,
  Traj02_Increased = ab,
  Traj02_Decrease = ac,

  
  Traj24_All = ba,
  Traj24_Increased = bb,
  Traj24_Decrease = bc,

  Traj45_All = ca,
  Traj45_Increased = cb,
  Traj45_Decrease = cc,
  
  Traj515_All = da,
  Traj515_ChimpPos = db,
  Traj515_Decrease = dc,

 
  Traj1530_All = ea,
  Traj1530_Increased = eb,
  Traj1530_Decrease = ec

)


cluster_plots <- lapply(names(cluster_lists), function(clust_name) {
  plot_cluster_fc_sig_trajectory(cluster_lists[[clust_name]], clust_name, TopTable_Traj)
})

names(cluster_plots) <- names(cluster_lists)

for (p in cluster_plots) print(p)


```

```{r Chimp_TE_Expression_Directionality_results}

direction_results <- readRDS("data/TEs/TE_Expression/Directions_chimp_tes.RDS")

direction_results <- lapply(direction_results, function(x) {
  
  x[c("all", "Increased", "Decreased")] <- lapply(
    x[c("all", "Increased", "Decreased")],
    function(df) {
      tidyr::separate(
        df,
        Gene,
        into = c("TE_Name", "TE_Family", "TE_Class"),
        sep = ":",
        remove = FALSE
      )
    }
  )
  
  x
})


direction_all <- lapply(direction_results, `[[`, "all")

direction_all <- dplyr::bind_rows(
  direction_all,
  .id = "Trajectory"
)

class_proportions_all <- direction_all %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions_all,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()

class_proportions <- direction_all %>%
  filter(direction == "Increased") %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of Increased TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()


class_proportions <- direction_all %>%
  filter(direction == "Decreased") %>%
  dplyr::count(Trajectory, TE_Class) %>%
  group_by(Trajectory) %>%
  mutate(Proportion = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Trajectory = factor(
      Trajectory,
      levels = c(
        "Traj02_genes",
        "Traj24_genes",
        "Traj45_genes",
        "Traj515_genes",
        "Traj1530_genes"
      )
    )
  )

ggplot(class_proportions,
       aes(x = Trajectory, y = Proportion, fill = TE_Class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Proportion of Decreased TE DEGs",
    fill = "TE Class"
  ) +
  theme_classic()
```


```{r Updating Website, eval = FALSE, fig.height=10, fig.width=10}

# git -> commit all changes
# git -> push
wflow_publish("analysis/TECount_Chimp.Rmd")

```
